---
title: "Profiling 5"
author: "mhv"
date: "12/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(scales)
library(tidyverse)

source("/Users/mhvpbp13/git/rsystrade/R/benchmark.R")
```

############################################
## See `Profiling-5-3.html`
############################################



# Test 2
## Define functions to be tested

```{r}
version1a <- function(df) {
  var1 <- 0
  var2 <- 0
  var3 <- 0

  v1af1 <- function(n) {var1 <<- df[n-1, 1]}
  v1af2 <- function(n) {var2 <<- df[n-1, 2] + log(var1)}
  v1af3 <- function(n) {var3 <<- df[n-1, 3] + log(var2)}

  for(n in 2:nrow(df)) {
    var1 <- v1af1(n)
    var2 <- v1af2(n)
    var3 <- v1af3(n)
    
    df[n, ] <- c(
      var1,
      var2,
      var3
    )
  }

  df
}

v6af1 <- function(row_n, prev_row) {
  prev_row[1]
}
v6af2 <- function(row_n, prev_row) {
  prev_row[2] + log(row_n[1])
}
v6af3 <- function(row_n, prev_row) {
  prev_row[3] + log(row_n[2])
}


version6a <- function(df) {
  force(df)
  row <- df[1, ]
  for(i in 2:nrow(df)) {
    prev_row <- df[i-1, ]
    row[1] <- v6af1(row, prev_row)
    row[2] <- v6af2(row, prev_row)
    row[3] <- v6af3(row, prev_row)

    df[i, ] <- row
  }
  df
}
```

## Run tests
```{r}
f_list2 <- list(
  version1 = "version1a(df)",
  version6 = "version6a(df)"
)
df_2 <- map_df(
  10^(2:3), run_benchmark, ncol = 3, functions = f_list2) %>% 
  order_graphs()
```

## Print test data
```{r}
plot_bm(df_2, "nrow")
```



## Test 3
## Define functions to be tested
This dynamically creates the function in version 1.
Of course this introduces various loops, so the two can not be compared directly.
```{r}
version1b <- function(df) {
  ncols <- ncol(df)
  for(m in 1:ncols) {
    assign(paste0("var", m), 0)
  }

  for(c in 2:ncols) {
    assign(paste0("v1bf", m), function(n) {
        assign(
          paste0("var", m),
          parse(
            text = paste0(
              "df[n-1, ", m, "] + log(var", c-1, ")"
            )
          ),
          envir = parent.frame()
        )
      }
    )
  }

  for(n in 2:nrow(df)) {
      for(m in 1:ncols) {
        assign(
          paste0("var", m), 
          parse(
            text = paste0("v1bf", m, "(", n, ")")
          )
        )
      }

    assign(
      paste0("df[", n, "]"),
      parse(
        text = paste0("var", m)
      )
    )

  }

  df
}

v6af1 <- function(row_n, prev_row) {
  prev_row[1]
}
v6af2 <- function(row_n, prev_row) {
  prev_row[2] + log(row_n[1])
}
v6af3 <- function(row_n, prev_row) {
  prev_row[3] + log(row_n[2])
}

v6bf <- function(row_n, prev_row, column) {
  prev_row[column]
}
version6b <- function(df) {
  force(df)
  ncols <- ncol(df)
  row <- df[1, ]
  for(n in 2:nrow(df)) {
    prev_row <- df[n-1, ]
    for(m in 1:ncols) {
      assign(
        paste0("row[", m, "]"),
        v6bf(row, prev_row, m)
      )
    }
    df[n, ] <- row
  }
  df
}
```

## Run tests: Compare dynamic and fixed number of columns
```{r}
#source("/Users/mhvpbp13/git/rsystrade/R/benchmark.R")
f_list3 <- list(
  version1a = "version1a(df)",
  version1b = "version1b(df)",
  version6a = "version6a(df)",
  version6b = "version6b(df)"
)
df_3_rows <- map_df(
    2^(7:11),
    ~ run_benchmark(
      nrow = .x,
      ncol = 3,
      functions = f_list3,
      times = 5,
      df = NA,
      elements = "same",
      element = 2
    )
  ) %>%
  order_graphs()
```

## Print test data
```{r}
plot_bm(df_3_rows, "nrow", mean_or_median = mean)
```
## Test 4

## Functions
Same as Test 3

## Run tests: 
### A) How does performance depend on number of rows?
### B) How does performance depend on number of columns?
```{r}
source("/Users/mhvpbp13/git/rsystrade/R/benchmark.R")
f_list3 <- list(
  version1b = "version1b(df)",
  version6b = "version6b(df)"
)

## A)
df_3_rows <- map_df(
    10^(2:4),
    ~ run_benchmark(
      nrow = .x,
      ncol = 3,
      functions = f_list3,
      times = 5,
      df = NA,
      elements = "same",
      element = 2
    )
  ) %>%
  order_graphs()

## B)
df_3_col <- map_df(
    2^(2:4),
    ~ run_benchmark(
      nrow = 100,
      ncol = .x,
      functions = f_list3,
      times = 5,
      df = NA,
      elements = "same",
      element = 2
    )
  ) %>%
  order_graphs()
```

## Print test data
### A) How does performance depend on number of rows?
```{r}
plot_bm(df_3_rows, "nrow", mean_or_median = mean)
```

### B) How does performance depend on number of columns?
```{r}
plot_bm(df_3_col, "ncol", mean_or_median = mean)
```


```{r}
#detach("package:microbenchmark", unload=TRUE)
#detach("package:ggplot2", unload=TRUE)
detach("package:tidyverse", unload=TRUE)
#detach("package:purrr", unload=TRUE)
#detach("package:tibble", unload=TRUE)
#detach("package:dplyr", unload=TRUE)
#detach("package:magrittr", unload=TRUE)
```
