---
title: "rsystrade"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rsystrade}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(rsystrade)
```

# About

*rsystrade* is a framework for researching, testing and executing systematic trading strategies.

# Philosophy

The intention is to provide as much transparency as possible about how things are calculated. So all functions that implement a formula live in files named `formulas-*.R`.

The intention is to *not* veil calculations in an endless maze of OOP abstractions and inheritance.

# The framework
## Using the framework


### Calculate position size

* $Z_C$, **position size** in *instrument contracts*.  
* $K_A$, trading **capital** in *account currency*.  
* $S$, **signal**.  
* $\sigma_{AST}$,  **annualized system risk** target in *decimal percentages*.  
* $\sigma_{AI}$, **annualized instrument risk** in *decimal percentages*.  
  * Measured as *daily price volatility*, which is the *volatility of daily returns* in decimal percentages.  
* $w$, **instrument weight**.  
* $\text{IDM}$, **Instrument Diversification Multiplier**.  
  * Measured as the value of a 1% increase in the price of a block.  
$$Z_C = S \cdot \frac{\sigma_{AST}}{\sigma_{AI}} \cdot K_A \cdot w \cdot \text{IDM}$$

## System components

A system consists of 

-   Any number of *algos*. An *algo* in this context is a distinct combination of a *trade rules pair variation* and data for an *instrument* (typically price data). For instance, a (20, 80) Moving Average Crossover plus a stop loss applied to a time series of S&P500 prices is one algo, and a (16, 64) MAC plus a stop loss applied to a time series of S&P500 prices is another algo. A (16, 64) MAC plus a stop loss applied to a time series of corn prices would be a third algo. Here a *rules pair* (or *rule* in short) is a combination of one *signal rule* and one optional *stop loss rule*.  

-   A `signal` table for each algo. The `signal` table contains the *signal* calculated by the algo, and some data produced by intermediate steps in the algo. A *signal* is the trade signal used to determine whether to enter, exit or modify a trade.  
-   A `position` table for each subsystem/instrument. All the different algos that affect a single instrument are combined to make up a single subsystem for that instrument. The *position* table contains the value of the resulting position for the subsystem as well as different data produced by intermediate steps.  
-   A `system_account` table containing aggregate data for the entire system.

When you create a new system, a list is created which contains the entire system. You create a new system with `make_system()`.
```r
my_system <- make_system(
    algos = algos,
    init_capital = 100000,
    risk_target = 0.12,
    risk_window_length = 25,
    stop_loss_fraction = 0.5,
    min_periods = 1,
    mode = "sim" ## "sim" or "live"
)
```
This list contains 

  * A list of instrument data sets. One data set for each unique instrument.  
  * The user provided list `algos`, which contains a list of algos, which in turn each contain an instrument name and a list containing one _signal rule_ and one optional _stop loss rule_.  
  * The list `signal_tables`, which contains a table as a data frame for each _algo_.  
  * The list `position_tables`, which contains a table as a data frame for each _instrument_.  
  * The data frame `system_account`, which contains a table as a data frame.
  

Once the `system` list has been build using `make_system()`, the system is updated with new data using `update_system`. This takes the entire `my_system` list as input and spits out an updated `my_system` list.
```r
update_system(my_system)
```


### Algos



### Signals



### Positions



### Accounts



## Simulation vs live trading

For each time interval `t` (typically daily basis) we get the new data (e.g. daily settlement price). For a simple *Moving Averages Crossover* rule, we would calculate the *moving averages* and the MAC signal for that day and append these to the data base. So the *moving averages* function actually only calculates *an* average for the given lookback period. It doesn't calculate *moving* averages for the entire time series. What makes the averages *move* is the daily calculations of averages.

# Data

## Simulation mode

Paths to data files are collected from the `algos` list. A list of unique data file paths is assembled. The structure of this list will be something like this:

``` r
datasets <- list(
  list(
    path = algos$algo1$datasets,
    data = read.csv(algos$algo1$datasets)
  ),
  list(
    path = algos$algo2$datasets,
    data = read.csv(algos$algo2$datasets)
  )
)
```

## Live mode

...

# Rules

Variations of the same rule are considered the separate rules. E.g. MA(16, 64) and MA(20, 80) would be two separate rules.

In rsystrade terminology we "enter" a trade rather than "open" a trade, and "exit" a trade rather than "close" a trade. This is simply to avoid any possible confusion with prices at *market open* and *market close*.


# Algos

An *algo* in **rsystrade** is a combination of a *data set* and a *rule*. The *data set* can be specified as an *instrument* (in `live` mode) or a reference to a *data file* (in `sim` mode).

# Signals

Signals are generated by rules. In **pysystemtrade**, these things are called *forecasts*. I call them signals, because they are *trade signals*, not forecasts. I reserve the term *forecast* for situations where we are making predictions.

### Scaled signals

> IMPORTANT!
>
> In **pysystemtrade**, *forecasts* are scaled such that the scaled signal has an expected absolute value of 10.
>
> In **rsystrade**, *signals* (forecasts) are by default scaled such that the scaled signal has an expected absolute value of 1. So if you would cap your signal within [-20, 20] in **pysystemtrade**, you should cap your signal within [-2, 2] in **rsystrade**.

An expected value of 1 seems intuitive. It is easy to relate to as a decimal ratio. A scale of 2 means double, or 200%. In the **pysystemtrade** framework you have to remember that a scale of 20 means double. But whatever works.

## Function names

**rsystrade** aspires to follow the convention, that function names should include a verb. However, because most of what **rsystrade** does is calculations, there is one exception to this guideline:

-   The name of an R function that implements a *math function* begins with `f_`.   
-   Here `f` at the beginning of function name represents the verb `calculate`.  
-   Instead of `calculate_normalization_factor()` we use `f_normalization_factor()`.  
-   In other words: Functions with a name that begin with `f_` perform *calculations*.  
-   As a mnemonic you might think of "formula", "function" or "find". I think of the math symbol for a function, because these calculations all implement some sort of formula. Visually the f at the beginning seems intuitive to me, when I skim the code, since it's an *R function* that implements a *math function*.

See also Guidelines.md.


# **rsystrade** terminology
## Algo
An *algo* in this context is a distinct combination of a *trade rules pair variation* and data for an *instrument* (typically price data). For instance, a (20, 80) Moving Average Crossover plus a stop loss applied to a time series of S&P500 prices is one algo, and a (16, 64) MAC plus a stop loss applied to a time series of S&P500 prices is another algo. A (16, 64) MAC plus a stop loss applied to a time series of corn prices would be a third algo. Here a *rules pair* is a combination of one entry rule and one optional exit rule.

## Entering/exiting (opening/closing)
We use *entering* and *exiting* rather than *opening* and *closing*, since "open" could indicate the price when the market opens, and "close" could indicate the price when the market closes.

## Rule
Aka a *trade rules pair*. A combination of one *entry rule* and one *exit rule*.

## Signal
A signal is an output number from a trade rule. The signal indicates whether we should enter or exit a position. The sign of the signal tells us if we should go long or short. The strength of the signal may indicate how big the position should be. Signals are proportional to expected risk adjusted returns. A signal of $0$ indicates that the position should be closed. The value of a signal is the product of the *entry rule signal* and the *exit rule signal*. The exit rule signal is $0$ to trigger exit (e.g. when a stop loss threshold is breached), and $1$ otherwise.

## Clamp (cap)
We use *clamp* rather than *cap* to indicate a restriction on a signal (or any value) which limits a signal to a lower and an upper limit. Typically a *cap* would be a restriction with an upper limit only. Sometimes a restriction with a lower limit only is called a *floor*. However, in mathematics *floor* means something else. The mathematical definition is implemented in the R function `floor()`. From the `?floor` help file: "`floor` takes a single numeric argument `x` and returns a numeric vector containing the largest integers not greater than the corresponding elements of `x`." To avoid confusion with the R function *floor* we don't use the term *floor* to indicate a lower restriction. Instead we say a *lower clamp*. Following this logic, we say *upper clamp* rather than *cap*.



