---
title: "Profiling 1"
author: "mhv"
date: "12/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library("microbenchmark")
library("ggplot2")
library("magrittr")
```


# Conclusion

+ Version 1:
  + Outer function `version1()` takes `df` as input.
  + Inner functions take no input.
    + Modify single elements in `df` in env of outer function based on other 
    single elements of `df` .
    + Each inner function is called in turn inside `version1()`.
+ Version 2: quasi-S3 OOP
  + Function `version2()` takes `df` as input.
  + Function `version2()` calls external functions which take `df` from env of
  `version2()`
  + External functions modify `df` in their own env and output modified `df`.
  + Each external function is called in turn inside `version2()` such that one 
  modifies the output of the previous one.
+ Version 3:
  + Same as version 2, except that the functions that modify `df` now are 
  defined inside the outer function `version3()`.
+ Version 4:
  + Like version 2, except that external functions are chained inside 
  `version4()` in a pipe with `%>%`.
+ Version 5: 
  + Functions output single values.
  + Change all modified elements in `df` in one step at the end.

There is no significant difference in performance between the different
approaches. In particular, there is no significant difference in performance
between the OOP (vserion 2) and FP (version 1+3+4) approaches.

Version 5 seems to show, that it's fine to store all variables in a large data frame. However, this turns not to be true, as seen in `Profiling 2.Rm` and `Profiling 3.Rmd`.

Question: Can we trust, that there is no difference between the approaches in
version 1 - 4? Test this...

Note: There is a big different between the first and second time we run the 
benchmark. What are the implications of this? Which situation applies most to
real world usage?

Looking at median to avoid leverage of outliers.

See also [/Users/mhvpbp13/Documents/Notes/Investering/rsystrade/Notes/rsystrade/Development/Coding paradigms - rsystrade.md](/Users/mhvpbp13/Documents/Notes/Investering/rsystrade/Notes/rsystrade/Development/Coding paradigms - rsystrade.md)

# Instructions:

Run everything at once.

# Performance test 1

```{r}
nrows <- 252 * 25 ## 6300 (25 years of trade data for single table)
#nrows <- 1000000
ncols <- 20
```


## Version 1
+ Outer function takes df as input
  + Inner functions take no input.
    + Modify single elements in df based on other single elements of df in
    env of outer function.

```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

version1 <- function(df) {
  v1f1 <- function() {
    df[1, 1] <<- "a"
  }
  v1f2 <- function() {
    df[1, 2] <<- paste0(df[1, 1], "b")
  }
  v1f3 <- function() {
    df[1, 3] <<- paste0(df[1, 2], "c")
  }
  v1f4 <- function() {
    df[1, 4] <<- paste0(df[1, 3], "d")
  }
  
  v1f1()
  v1f2()
  v1f3()
  v1f4()
 
  df
}

version1_out <- version1(df)
version1_out[1, 1:4]

rm(df)
```

## Version 2: Quasi-S3 OOP
The functions `v5f*()` can be thought of as methods that operate on the data frame df, and the function `version5()` can be thought of as a constructor that creates a new object (the modified data frame) by applying these methods in a specific order.
```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

v2f1 <- function(df) {
  df[1, 1] <- "a"
  df
}
v2f2 <- function(df) {
  df[1, 2] <- paste0(df[1, 1], "b")
  df
}
v2f3 <- function(df) {
  df[1, 3] <- paste0(df[1, 2], "c")
  df
}
v2f4 <- function(df) {
  df[1, 4] <- paste0(df[1, 3], "d")
  df
}

version2 <- function(df) {
  df <- v2f1(df)
  df <- v2f2(df)
  df <- v2f3(df)
  df <- v2f4(df)
  
  df
}

version2_out <- version2(df)
version2_out[1, 1:4]

rm(df)
```

## Version 3

```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

version3 <- function(df) {
  v3f1 <- function(df) {
    df[1, 1] <- "a"
    df
  }
  v3f2 <- function(df) {
    df[1, 2] <- paste0(df[1, 1], "b")
    df
  }
  v3f3 <- function(df) {
    df[1, 3] <- paste0(df[1, 2], "c")
    df
  }
  v3f4 <- function(df) {
    df[1, 4] <- paste0(df[1, 3], "d")
    df
  }
  
  df <- v3f1(df)
  df <- v3f2(df)
  df <- v3f3(df)
  df <- v3f4(df)
  
  df
}

version3_out <- version3(df)
version3_out[1, 1:4]

rm(df)
```


## Version 4

```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

v4f1 <- function(df) {
  df[1, 1] <- "a"
  df
}

v4f2 <- function(df) {
  df[1, 2] <- paste0(df[1, 1], "b")
  df
}

v4f3 <- function(df) {
  df[1, 3] <- paste0(df[1, 2], "c")
  df
}

v4f4 <- function(df) {
  df[1, 4] <- paste0(df[1, 3], "d")
  df
}

version4 <- function(df) {
  df %>%
    v4f1 %>%
    v4f2 %>%
    v4f3 %>%
    v4f4
}

version4_out <- version4(df)
version4_out[1, 1:4]

rm(df)
```



## Version 5
Functions output single values.  
Change all modified elements in df in one step at the end.

```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

version5 <- function(df) {
  v5f1 <- function() {
    "a"
  }
  v5f2 <- function(prev) {
    paste0(prev, "b")
  }
  v5f3 <- function(prev) {
    paste0(prev, "c")
  }
  v5f4 <- function(prev) {
    paste0(prev, "d")
  }
  
  out1 <- v5f1()
  out2 <- v5f2(out1)
  out3 <- v5f3(out2)
  out4 <- v5f4(out3)
  
  df[1, 1:4] <- c(out1, out2, out3, out4)
  
  df
}

version5_out <- version5(df)
version5_out[1, 1:4]

rm(df)
```


```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

mb_out <- microbenchmark(
  version1(df),
  version2(df),
  version3(df),
  version4(df),
  version5(df),
  times = 20L
)

mb_plot <- autoplot(mb_out) +
  geom_jitter(position = position_jitter(0.2, 0), 
              aes(color = expr), alpha = 0.4) + 
  aes(fill = I("gray")) + 
  theme(legend.position = "none") 
mb_plot
#labs(title = "Kernel smoother", subtitle = "")

rm(df)
```

```{r}
summary(mb_out)
```

Run benchmark test again
```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

mb_out <- microbenchmark(
  version1(df),
  version2(df),
  version3(df),
  version4(df),
  version5(df),
  times = 20L
)
summary(mb_out)

rm(df)
```









```{r eval=FALSE, include=FALSE}
detach("package:microbenchmark", unload=TRUE)
detach("package:ggplot2", unload=TRUE)
detach("package:purrr", unload=TRUE)
detach("package:tibble", unload=TRUE)
detach("package:dplyr", unload=TRUE)
detach("package:magrittr", unload=TRUE)
```


