---
title: "Bootstrap dev"
author: "mhv"
date: "2025-05-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Bootstrapping

## Basic moving block bootstrap

*This bootstrap works with dependent data, however, the bootstrapped observations will not be stationary anymore by construction. But, it was shown that varying randomly the block length can avoid this problem.[30] This method is known as the stationary bootstrap. Other related modifications of the moving block bootstrap are the Markovian bootstrap and a stationary bootstrap method that matches subsequent blocks based on standard deviation matching.*
https://en.wikipedia.org/wiki/Bootstrapping_(statistics)#Block_bootstrap 
```{r}
#' x timeseries or vector to sample from
#' block_length block length
#' func function to apply to sample
#' num_blocks number of blocks per bootstrap
#' num_boots number of bootstrap series
rsbootstrap <- function(x, block_length = 1, func = mean, num_blocks, num_boots) {
  n <- length(x)
  func_out <- NULL
  for(bs in 1:num_boots) {
    ids <- sample(1:(n - block_length + 1), num_blocks, replace = TRUE)
    if(block_length > 1) {
      ids_tmp <- NULL
      for(i in seq_along(ids)) {
        ids_tmp <- c(ids_tmp, (ids[i]):(ids[i] + block_length - 1))
      }
      ids <- ids_tmp
    }
    func_out <- c(func_out, func(x[ids]))
  }
  mean(func_out)
}
```

```{r}
rsbootstrap(x = rnorm(500), block_length = 5, statistic = sd, num_blocks = 20, num_boots = 10000)
```

## Selecting block length

```{r}
set.seed(1729)
#x_data <- ts(100 * c(1, cumprod(1 + rnorm(500, 0, 0.01))), start = c(1970,1,1), frequency = 12)
x_data <- ts(rnorm(500, 0, 0.05), start = c(1970,1,1), frequency = 12)
```

Let's try to detrend the data so that we can fit an ARIMA....
```{r}
x_data_decomp <- decompose(x_data)
plot(x_data_decomp)
```

Detrended
```{r}

plot(x_data, ylim = c(min(c(x_data, x_data - x_data_decomp$trend), na.rm = TRUE), max(c(x_data, x_data - x_data_decomp$trend), na.rm = TRUE)))
lines(x_data - x_data_decomp$trend, col = "red")
```


How do autocorrelations look?

```{r}
acf(x_data, lag.max = 500)
```

```{r}
pacf(x_data, lag.max = 20)
```



*For a white noise series, we expect 95% of the spikes in the ACF to lie within ±2/√T where T is the length of the time series. It is common to plot these bounds on a graph of the ACF (the blue dashed lines above). If one or more large spikes are outside these bounds, or if substantially more than 5% of spikes are outside these bounds, then the series is probably not white noise.*
https://otexts.com/fpp2/wn.html#wn 
```{r}
threshold <- 2/(sqrt(500))
```

Here we select the first lag that drops below the threshold to determine the length of a block.  
This method is problematic, because a white noise series will damp very slowly, if at all. So we might end up with the entire time series being one block.
```{r}
find_optimal_block_length <- function(x, threshold) {
  ac <- acf(x_data, lag.max = 100, plot = FALSE)
  min(which(ac$acf < threshold))
}
```


```{r}
block_length <- find_optimal_block_length(x_data, 2/(sqrt(500)))
num_blocks <- round(length(x_data)/(10 * block_length))
rsbootstrap(
  x = x_data, 
  block_length = block_length, 
  statistic = sd, 
  num_blocks = num_blocks, 
  num_boots = 10000
)
```


## `tsbootstrap`

Alternatively we could depend on the `tseries` package
```{r}
library(tseries)
```

This, however, does not free us from picking a block length, or rather in this case an average block length.
```{r}
tsbootstrap_output <- tsbootstrap(
  x_data, 
  nb = 1000, 
  statistic = mean, 
  m = 1, 
  b =25,
  type = "stationary"
)
mean(tsbootstrap_output$statistic)
```

Compare to mean of input `x_data`
```{r}
mean(x_data)
```

Which is way better than fixed block length.

## Multivariate block bootstrap

```{r}
#' x data frame
#' block_length block length
#' func function to apply to sample
#' apply_func_to c("df", "cols") apply function to entire data frame 
#'   or column by column
#' num_blocks number of blocks per bootstrap
#' num_boots number of bootstrap series
rsbootstrap <- function(
    x, 
    block_length = 1, 
    func, 
    apply_func_to,
    num_blocks, 
    num_boots) {
  n <- nrow(x)
  func_out <- NULL
  
  make_blocks <- function(ids) {
    ids_tmp <- NULL
    for(i in seq_along(ids)) {
        ids_tmp <- c(ids_tmp, (ids[i]):(ids[i] + block_length - 1))
    }
    ids_tmp
  }
  
  make_ids <- function() {
    ids <- sample(1:(n - block_length + 1), num_blocks, replace = TRUE)
    if(block_length > 1) {
      ids <- make_blocks(ids)
    }
    ids
  }

  make_bootstrap_series <- function(x) {
    as.data.frame(lapply(x, function(column) {column[make_ids()]}))
  }
  
  ## TODO: Impement with replicate() instead of loops
  bootstrap_series <- list()
  func_outs <- list()
  
  if(apply_func_to == "cols") {
    for(i in 1:num_boots) {
      bootstrap_series[[i]] <- make_bootstrap_series(x)
      func_outs[[i]] <- as.data.frame(lapply(bootstrap_series[[i]], function(x) {func(x)}))
    }
  } else if(apply_func_to == "df") {
    for(i in 1:num_boots) {
      bootstrap_series[[i]] <- make_bootstrap_series(x)
      func_outs[[i]] <- func(bootstrap_series[[i]])
    }
  } else (warning("apply_func_to takes values \"cols\" or \"df\"."))
  
  list(
    bootstrap_series = bootstrap_series,
    func_outs = func_outs,
    func_out_means = lapply(func_outs, function(x) {mean(unlist(x), na.rm = TRUE)})
  )
}
```


test:
```{r}
set.seed(1729)
x_df <- data.frame(a = rnorm(100), b = rnorm(100), c = rnorm(100))

bs_out <- rsbootstrap(
  x_df, 
  block_length = 10, 
  func = mean, 
  apply_func_to = "cols",
  num_blocks = 20, 
  num_boots = 30
)
```

```{r}
set.seed(1729)
x_df <- data.frame(a = rnorm(100), b = rnorm(100), c = rnorm(100))
test_func <- function(x) {lapply(x, sd)}

bs_out <- rsbootstrap(
  x_df, 
  block_length = 10, 
  func = test_func, 
  apply_func_to = "df",
  num_blocks = 20, 
  num_boots = 30
)
```

## Own implementation

### Functions

NOTE: See also blockBootstrap.cpp

```{r}
Rcpp::sourceCpp("/Users/mhvpbp13/git/rsystrade/cpp/blockBootstrap.cpp") 
```


```{r}

# Required packages
library(Rcpp)         # For high-performance bootstrap sampling
library(zoo)          # For time series handling and lag operations
library(parallel)     # For multi-core parallel execution
library(depmixS4)     # For fitting Hidden Markov Models (HMM)
library(dplyr)        # For tidy data manipulation (e.g., lagging)
library(MSwM)         # For Markov-Switching AR models (MS-AR)


# HMM-based block identification and sampling bootstrap
hmm_bootstrap <- function(returns, nstates = 2, num_blocks = 100, num_boots = 100) {
  df <- data.frame(return = returns)

  # Fit HMM to identify regimes
  model <- depmix(return ~ 1, data = df, nstates = nstates, family = gaussian())
  fit <- fit(model, verbose = FALSE)
  states <- posterior(fit, type = "viterbi")$state

  # Identify continuous blocks of the same regime
  r <- rle(states)
  ends <- cumsum(r$lengths)
  starts <- c(1, head(ends, -1) + 1)  # Starts one after previous end
  blocks <- Map(function(s, e, st) list(start = s, end = e, state = st), starts, ends, r$values)

  # Separate blocks into trend and noise (assuming trend is state 1)
  trend_blocks <- Filter(function(b) b$state == 1, blocks)
  noise_blocks <- Filter(function(b) b$state != 1, blocks)

  # Function to extract a block from returns
  get_block <- function(b) returns[b$start:b$end]

  # Sample trend and noise blocks to build bootstrapped series
  sampled_series <- replicate(num_boots, {
    sampled_trend <- lapply(sample(trend_blocks, num_blocks, replace = TRUE), get_block)
    sampled_noise <- lapply(sample(noise_blocks, num_blocks, replace = TRUE), get_block)
    unlist(c(sampled_trend, sampled_noise))
  }, simplify = FALSE)

  return(sampled_series)
}


# Markov-switching AR-based block bootstrap (nonstationary block identification)

#' MSAR-Based Bootstrap for Time Series
#'
#' Performs a bootstrap using blocks identified by a Markov-switching autoregressive (MSAR) model.
#' The time series is segmented into regimes (states) based on an MSAR model, and blocks are sampled
#' from the same inferred regime to preserve regime-dependent dynamics.
#'
#' @param x A numeric vector representing a univariate time series.
#' @param ar_order Integer specifying the order of the autoregressive model for each state.
#' @param nstates Number of regimes (states) in the MSAR model.
#' @param num_blocks Number of blocks to sample for each bootstrap replicate.
#' @param num_boots Number of bootstrap replicates to generate.
#'
#' @return A list of bootstrap replicate time series, each as a numeric vector of the same length as \code{num_blocks * block_length}.
#'
#' @details This function fits an MSAR model using the \code{MSwM} package and samples blocks
#' from the time series corresponding to segments assigned to the same state. This allows capturing
#' non-stationary dynamics that follow regime-switching patterns.
#'
#' Note: Block lengths are determined by the contiguous segments within the same state, not by a fixed value.
#'
#' @examples
#' \dontrun{
#' set.seed(123)
#' x <- arima.sim(n = 200, list(ar = 0.8))
#' msar_samples <- msar_bootstrap(x, ar_order = 1, nstates = 2, num_blocks = 4, num_boots = 10)
#' plot(msar_samples[[1]], type = "l")
#' }
#'
#' @importFrom MSwM msmFit
#' @importFrom stats lm ar rnorm
#' @export
msar_bootstrap <- function(returns, ar_order = 1, nstates = 2, num_blocks = 100, num_boots = 100) {
  df <- data.frame(y = returns)
  for (i in 1:ar_order) df[[paste0("lag", i)]] <- dplyr::lag(returns, i)
  df <- na.omit(df)  # Remove initial NA rows

  base_model <- lm(y ~ ., data = df)  # Fit base AR(p) model
  ms_model <- msmFit(base_model, k = nstates, sw = rep(TRUE, length(coef(base_model)) + 1))  # Switch regime per coefficient

  states <- ms_model@Fit@smoProb
  state_seq <- apply(states, 1, which.max)  # Most probable state per time step

  r <- rle(state_seq)  # Run-length encode states
  ends <- cumsum(r$lengths)
  starts <- c(1, head(ends, -1) + 1)
  blocks <- Map(function(s, e, st) list(start = s, end = e, state = st), starts, ends, r$values)

  # Separate trend vs. noise regimes (e.g., regime 1 as trend)
  trend_blocks <- Filter(function(b) b$state == 1, blocks)
  noise_blocks <- Filter(function(b) b$state != 1, blocks)

  get_block <- function(b) returns[b$start:b$end]

  sampled_series <- replicate(num_boots, {
    sampled_trend <- lapply(sample(trend_blocks, num_blocks, replace = TRUE), get_block)
    sampled_noise <- lapply(sample(noise_blocks, num_blocks, replace = TRUE), get_block)
    unlist(c(sampled_trend, sampled_noise))
  }, simplify = FALSE)

  return(sampled_series)
}

#' Perform Wild Bootstrap for regression residual-like data
#'
#' @param x Numeric vector or matrix (time series or residuals).
#' @param num_boots Number of bootstrap replicates.
#' @return List of bootstrap samples
wild_bootstrap <- function(x, num_boots) {
  if (is.vector(x)) {
    x <- matrix(x, ncol = 1)
  }
  n <- nrow(x)
  d <- ncol(x)
  
  bootstrap_samples <- vector("list", num_boots)
  
  for (b in seq_len(num_boots)) {
    # Generate Rademacher weights: +1 or -1 with probability 0.5
    v <- sample(c(-1, 1), size = n, replace = TRUE)
    # Multiply each row (observation) by corresponding weight
    boot_sample <- x * v
    bootstrap_samples[[b]] <- boot_sample
  }
  bootstrap_samples
}

# Time-series-aware k-fold cross-validation to estimate optimal p (for stationary bootstrap)
k_fold_cv_ts <- function(x, y, k = 5, p_grid = seq(0.01, 0.5, by = 0.01), model_func, score_func) {
  n <- nrow(x)
  fold_size <- floor(n / k)  # Ensures consistent fold size across all folds (important for TS continuity)
  ends <- seq(fold_size, by = fold_size, length.out = k)  # Ending indices for each fold
  starts <- c(1, head(ends, -1) + 1)  # Starting indices ensure contiguous folds

  results <- sapply(p_grid, function(p_val) {
    scores <- mapply(function(start, end) {
      val_idx <- start:end
      train_idx <- setdiff(1:n, val_idx)

      x_train <- x[train_idx, , drop = FALSE]
      y_train <- y[train_idx]
      x_val   <- x[val_idx, , drop = FALSE]
      y_val   <- y[val_idx]

      model <- model_func(x_train, y_train)
      preds <- predict(model, newdata = as.data.frame(x_val))
      score_func(preds, y_val)
    }, starts, ends)
    mean(unlist(scores))  # Mean score for current p
  })

  best_p <- p_grid[which.min(results)]  # p minimizing error
  return(best_p)
}

# Mean Square Error score function
mse <- function(pred, actual) {
  mean((pred - actual)^2)
}

# Mean Absolute Error score function
mae <- function(pred, actual) {
  mean(abs(pred - true))
}


  
# General bootstrap function: wraps all block types and plug-in/cv-based p selection

#' Bootstrap Time Series with Flexible Block Structures
#'
#' Perform time-series-aware block bootstrap using moving, stationary,
#' or Markov-switching AR (MSAR) block schemes. Includes automatic
#' estimation of average block length via cross-validation or plugin method.
#'
#' @param x A numeric vector, matrix, or data frame of time series data.
#' @param n_length Length of bootstraped series. If num_blocks > 0, n_length is
#'   overwritten and calculated as num_blocks * block_length
#' @param block_length Length of individual blocks (ignored for MSAR).
#' @param block_type Type of block sampling: "moving", "stationary", "hmm", 
#'   "msar", or "wild".
#' @param func A function to apply to each bootstrap sample (e.g., mean).
#' @param apply_func_to Apply function to "cols" (columns individually) or "df" (whole data.frame).
#' @param num_blocks Number of blocks per bootstrap sample.
#' @param num_boots Number of bootstrap replicates.
#' @param p_method Method to estimate average block probability p: "1/n",
#'   "plugin", or "cross validation".
#' @param p Optional value for p. If NULL, uses p_method to estimate.
#' @param overlap Whether to allow overlapping blocks (for moving/stationary).
#' @param ar_order Order of AR model used in MSAR block segmentation.
#' @param nstates Number of states in the MSAR model.
#' @param model_func Model function used when p_method = "cross validation".
#' @param score_func Scoring function used when p_method = "cross validation". 
#'   Default: "mse"
#'
#' @return A list with:
#' \\describe{
#'   \\item{bootstrap_series}{List of sampled time series.}
#'   \\item{func_outs}{List of outputs from applying func to each sample.}
#'   \\item{func_out_means}{Mean output across bootstrap samples.}
#' }
#'
#' @examples
#' set.seed(42)
#' x <- arima.sim(n = 100, list(ar = 0.8))
#' result <- bootstrap(
#'   x = x,
#'   block_length = 10,
#'   block_type = "stationary",
#'   num_blocks = 5,
#'   num_boots = 100,
#'   func = mean,
#'   apply_func_to = "cols",
#'   p_method = "plugin"
#' )
#' print(result$func_out_means)
#'
#' @importFrom stats acf predict
#' @importFrom utils head
#' @export
bootstrap <- function(
    x, 
    n_length = 100,
    block_length = 10, 
    block_type = "stationary",
    func = mean, 
    apply_func_to = "cols",
    num_blocks = 10, 
    num_boots = 100, 
    p_method = "1/n",
    p = NULL, 
    overlap = TRUE, 
    ar_order = 1, 
    nstates = 2,
    model_func = NULL, 
    score_func = NULL) {

  block_type <- match.arg(block_type, choices = c("moving", "stationary", "hmm", "msar", "wild"))
  apply_func_to <- match.arg(apply_func_to, choices = c("cols", "df"))
  p_method <- match.arg(p_method, choices = c("1/n", "cross validation", "plugin"))

  if (is.vector(x)) x <- as.matrix(x)
  if (is.data.frame(x)) x <- as.matrix(x)

  n <- nrow(x)
  d <- ncol(x)

  if (block_type == "stationary" && is.null(p)) {
    if (p_method == "1/n") {
      p <- 1 / block_length  # Simple heuristic: average block length = block_length
    } else if (p_method == "plugin") {
      ac1 <- acf(x, lag.max = 1, plot = FALSE)$acf[2,,1]  # First-lag autocorrelation
      p <- 1 - abs(mean(ac1, na.rm = TRUE))  # High autocorrelation -> smaller p
    } else if (p_method == "cross validation") {
      if (is.null(model_func) || is.null(score_func)) {
        stop("For cross validation, provide model_func and score_func.")
      }
      y <- x[, 1]  # Assume first column is target for model
      p <- k_fold_cv_ts(x, y, model_func = model_func, score_func = score_func)
    }
  }

  bootstrap_series <-
    if (block_type %in% c("moving", "stationary")) {
      blockBootstrap(x, n_length, block_length, num_blocks, num_boots, block_type, p, overlap)
    } else if (block_type == "hmm") {
      lapply(hmm_bootstrap(x[, 1], nstates = nstates,
                            num_blocks = num_blocks, num_boots = num_boots),
             function(series) matrix(series, ncol = 1))
    } else if (block_type == "msar") {
      lapply(msar_bootstrap(x[, 1], ar_order = ar_order, nstates = nstates,
                            num_blocks = num_blocks, num_boots = num_boots),
             function(series) matrix(series, ncol = 1))
    } else if (block_type == "wild") {
      wild_bootstrap(x, num_boots)
    } else {
      stop("Unsupported block_type.")
    }

  func_outs <- lapply(bootstrap_series, function(sampled) {
    if (apply_func_to == "df") {
      func(as.data.frame(sampled))
    } else {
      apply(sampled, 2, func)
    }
  })

  func_out_means <- Reduce("+", func_outs) / length(func_outs)  # Vectorized aggregation

  return(list(
    bootstrap_series = bootstrap_series,
    func_outs = func_outs,
    func_out_means = func_out_means
  ))
}
```


### Tests

#### Inspect

Compare `rst()` and `rsstd()`
```{r}
## xi: location (mean)
## omega: scale (sd)
## alpha: slant (skewness)
## nu: df
x_rst <- rst(1000, xi = 0, omega = 1, alpha = 5, nu = 3)

## mean: location (mean)
## sd: scale (sd)
## xi: slant (skewness)
## nu: df
x_sstd <- fGarch::rsstd(1000, mean = 0, sd = 1, xi = 5, nu = 3)

test_df <-  data.frame(x_rst = x_rst, x_sstd = x_sstd) %>% 
  gather(key = "dist", value = "value")

# Violin plots
ggplot(test_df, aes(x = dist, y = value, fill = dist)) +
  geom_violin() + geom_jitter(width = 0.1, alpha = 0.3) +
  theme_minimal()
```



```{r}
# Load required packages
library(sn)       # For skewed t-distribution
library(ggplot2)  # For visualization

# Simulate skewed t-distributed AR(1) returns
set.seed(123)
n <- 300
phi <- 0.6
nu <- 3
alpha <- 5
innov <- rst(n, xi = 0, omega = 1, alpha = alpha, nu = nu)

returns <- numeric(n)
returns[1] <- innov[1]
for (t in 2:n) {
  returns[t] <- phi * returns[t - 1] + innov[t]
}

# Bootstrap variants using your unified bootstrap() function
b1 <- bootstrap(returns, block_type = "moving", num_boots = 1)[[1]][[1]]
b2 <- bootstrap(returns, block_type = "stationary", p_method = "cross validation", num_boots = 1)[[1]][[1]]
b3 <- bootstrap(returns, block_type = "wild", num_boots = 1)[[1]][[1]]
b4 <- bootstrap(returns, block_type = "hmm", nstates = 2, num_boots = 1)[[1]][[1]]
b5 <- bootstrap(returns, block_type = "msar", ar_order = 1, nstates = 2, num_boots = 1)[[1]][[1]]

# Helper: convert to data.frame with id and time
to_df <- function(series, label) {
  data.frame(value = series, 
             time = seq_along(series), 
             type = label)
}

# Combine
df_plot <- bind_rows(
  to_df(returns, "original"),
  to_df(b1, "wild"),
  to_df(b2, "stationary"),
  to_df(b3, "moving"),
  to_df(b4, "hmm"),
  to_df(b5, "msar")
)

# Plot
ggplot(df_plot, aes(x = time, y = value)) +
  geom_line() +
  facet_wrap(~type, scales = "free_y", ncol = 2) +
  theme_minimal() +
  labs(title = "Bootstrap Samples from Different Methods")
```

#### Setup: Simulate returns
```{r}
#library(sn)       # For skewed t-distribution
library(e1071)    # For skewness, kurtosis
library(ggplot2)
library(dplyr)
library("fGarch") # For skewed t-distribution
```


```{r}
# Simulate skewed t-distributed annual log-returns
n_years = 1

n <- 252 * n_years
params <- list(m = 0.052, s = 0.098, nu = 3.032, xi = 0.023)

set.seed(1729)
returns <- rsstd(n, params$m, params$s, params$nu, params$xi)
```



#### Run bootstraps
```{r}
boot_methods <- c("moving", "stationary", "wild", "hmm", "msar")
n_boots <- 10000

boots <- lapply(boot_methods, function(m) {
  bootstrap(matrix(returns, ncol = 1),
            block_type = m,
            num_boots = n_boots,
            block_length = 20,
            overlap = TRUE,
            nstates = 2,
            ar_order = 1,
            p_method = "1/n")
})
names(boots) <- boot_methods
```

#### Compare bootstrapped moments
```{r}
# Compute summary stats
summary_stats <- function(b) {
  sapply(b, function(x) {
    x <- as.numeric(x[[1]])
    c(mean = mean(x), skew = skewness(x), kurt = kurtosis(x))
  }) %>% t() %>% as.data.frame()
}

moments_df <- bind_rows(
  lapply(seq_along(boots), function(i) {
    df <- summary_stats(boots[[i]])
    df$method <- boot_methods[i]
    df
  })
)
```


```{r}
# Violin plots
ggplot(moments_df, aes(x = method, y = mean, fill = method)) +
  geom_violin() + geom_jitter(width = 0.1, alpha = 0.3) +
  theme_minimal() + ggtitle("Bootstrapped Mean Distributions")
```

```{r}
ggplot(moments_df, aes(x = method, y = skew, fill = method)) +
  geom_violin() + geom_jitter(width = 0.1, alpha = 0.3) +
  theme_minimal() + ggtitle("Bootstrapped Skewness Distributions")
```

```{r}
ggplot(moments_df, aes(x = method, y = kurt, fill = method)) +
  geom_violin() + geom_jitter(width = 0.1, alpha = 0.3) +
  theme_minimal() + ggtitle("Bootstrapped Kurtosis Distributions")
```


#### Fan plot (bootstrap trajectories)
```{r}
method <- "stationary"
mat <- do.call(cbind, boots[[method]]$bootstrap_series) 

# Compute quantiles across bootstrap replicates
quant_df <- apply(mat, 1, function(row) {
  quantile(row, probs = c(0.05, 0.25, 0.5, 0.75, 0.95))
}) %>%
  t() %>%
  as.data.frame() %>%
  mutate(time = 1:nrow(.))

names(quant_df)[1:5] <- c("q05", "q25", "q50", "q75", "q95")

# Plot fan chart
ggplot(quant_df, aes(x = time)) +
  geom_ribbon(aes(ymin = q05, ymax = q95), fill = "skyblue", alpha = 0.2) +
  geom_ribbon(aes(ymin = q25, ymax = q75), fill = "skyblue", alpha = 0.4) +
  geom_line(aes(y = q50), color = "black") +
  labs(title = "Fan Plot of Bootstrap Replicates",
       x = "Time", y = "Value") +
  theme_minimal()
```

#### Comparison to Theoretical Moments
For a standardized skewed t-distribution with 3 degrees of freedom and skewness parameter α ≈ 0.9:

- Mean ≈ 0 (after standardization)
- Skewness depends on α, but is > 0
- Kurtosis for t(3) is ≈ 9 (excess kurtosis ≈ 6)

```{r}
# Helper to get named moment vector
get_moments <- function(x) {
  c(
    mean = mean(x),
    variance = var(x),
    skewness = e1071::skewness(x),
    kurtosis = e1071::kurtosis(x)
  )
}

# Compute average moments across bootstrapped series
moment_list <- lapply(boot_methods, function(method) {
  boot_series <- boots[[method]]$bootstrap_series
  moment_matrix <- sapply(boot_series, get_moments)  # 4 x n_boots matrix
  rowMeans(moment_matrix)  # average across bootstraps
})

# Add theoretical skewed-t moments (mean = 0, variance = 3, skew = 0, kurt = Inf)
moment_matrix <- do.call(cbind, moment_list)

# Add theoretical values as a new column, not a row
moment_df <- as.data.frame(moment_matrix)
rownames(moment_df) <- c("mean", "variance", "skewness", "kurtosis")
colnames(moment_df) <- boot_methods
kurtosis_rsstd <- function(nu, xi) {
  if (nu <= 4) return(NA)  # Kurtosis undefined for ν ≤ 4
  num <- (xi^5 + xi^(-5)) * (xi + xi^(-1))
  den <- (xi^3 + xi^(-3))^2
  3 + (6 / (nu - 4)) * num / den
}

moment_df <- cbind(Theoretical = c(params$m, params$s^2, params$xi, kurtosis_sstd(params$nu, params$xi)), moment_df)

# Round for presentation
round(moment_df, 3)
```


Note: The Wild Bootstrap doesn't capture skewness by design.



```{r}
# Theoretical values (approximate, for comparison only)
true_mean <- 0
true_skew <- skewness(rsstd(100000, params$m, params$s, params$nu, params$xi))
true_kurt <- kurtosis(rsstd(100000, params$m, params$s, params$nu, params$xi))

plot_moment <- function(moment) {
  g <- ggplot(moments_df, aes_string(x = "method", y = moment, fill = "method")) +
    geom_violin() +
    geom_jitter(width = 0.1, alpha = 0.3) +
    geom_hline(yintercept = get(paste0("true_", moment)), linetype = "dashed", color = "red") +
    theme_minimal() +
    ggtitle(paste("Bootstrapped", toupper(moment), "vs Theoretical"))
  print(g)
}
```

```{r}
plot_moment("mean")
```

```{r}
plot_moment("skew")
```

```{r}
plot_moment("kurt")
```


#### Heatmap of Pairwise Correlations Across Bootstraps
```{r}
library(corrplot)

plot_correlation_heatmap <- function(method, B = 20) {
  method <- "stationary"
  mat <- do.call(cbind, boots[[method]]$bootstrap_series[1:B]) 
  cor_mat <- cor(mat)
  corrplot::corrplot(cor_mat, method = "color", type = "lower",
                     title = paste("Correlation across", B, method, "bootstraps"),
                     tl.cex = 0.8)
}

# Example for stationary bootstrap
plot_correlation_heatmap("stationary")
```



  
  
### Estimate skew t-distribution parameters

Simulate skewed t-distributed annual log-returns
```{r}
n_years = 10

n <- 252 * n_years
params <- list(m = 0.052, s = 0.098, nu = 3.032, xi = 0.023)

set.seed(1729)
returns <- rsstd(n, params$m, params$s, params$nu, params$xi)
```

Functions
```{r}
## Fit data to a skewed t, standardized t or normal distribution
## distribution = "sstd", "std" or "normal"
## If "normal", method="Nelder-Mead" is recommended, else method="BFGS"
fit_distribution <- function(x, method = "BFGS", distribution = "sstd") {
  
  if(distribution == "sstd") {
    loglik_sstd = function(beta) sum(- dsstd(x, mean = beta[1], sd = beta[2], nu = beta[3], xi = beta[4], log = TRUE))
    start = c(mean(x), sd(x), 3, 1)
  } else if(distribution == "std") {
    loglik_sstd = function(beta) sum(- dsstd(x, mean = beta[1], sd = beta[2], nu = beta[3], log = TRUE))
    start = c(mean(x), sd(x), 3)
  } else {
    loglik_sstd = function(beta) sum(- dnorm(x, mean = beta[1], sd = beta[2], log = TRUE))
    start = c(mean(x), sd(x))
  }
  
  optim_out = optim(start, loglik_sstd, method = method)
  
  n = length(x)
  
  if(distribution == "sstd") {
    fit <- qsstd((1:n - 0.5)/n, mean = optim_out$par[1], sd = optim_out$par[2], nu = optim_out$par[3], xi = optim_out$par[4])
    quantile_data <- qsstd(seq(0.005, 0.995, length.out = 600), mean = optim_out$par[1], sd = optim_out$par[2], nu = optim_out$par[3], xi = optim_out$par[4])
    dist_data <- psstd(seq(-0.3, 0.3, length.out = 600), mean = optim_out$par[1], sd = optim_out$par[2], nu = optim_out$par[3], xi = optim_out$par[4])
    dens_data <- dsstd(seq(-0.3, 0.3, length.out = 600), mean = optim_out$par[1], sd = optim_out$par[2], nu = optim_out$par[3], xi = optim_out$par[4])
    theoretical_quantiles <- qsstd(ppoints(n),optim_out$par[1], optim_out$par[2], optim_out$par[3], optim_out$par[4])
  } else if (distribution == "std") {
    fit <- qstd((1:n - 0.5)/n, mean = optim_out$par[1], sd = optim_out$par[2], nu = optim_out$par[3])
    quantile_data <- qstd(seq(0.005, 0.995, length.out = 600), mean = optim_out$par[1], sd = optim_out$par[2], nu = optim_out$par[3])
    dist_data <- pstd(seq(-0.3, 0.3, length.out = 600), mean = optim_out$par[1], sd = optim_out$par[2], nu = optim_out$par[3])
    dens_data <- dstd(seq(-0.3, 0.3, length.out = 600), mean = optim_out$par[1], sd = optim_out$par[2], nu = optim_out$par[3])
    theoretical_quantiles <- qstd(ppoints(n), mean = optim_out$par[1], sd = optim_out$par[2], nu = optim_out$par[3])
  } else {
    fit <- qnorm((1:n - 0.5)/n, mean = optim_out$par[1], sd = optim_out$par[2])
    quantile_data <- qnorm(seq(0.005, 0.995, length.out = 600), mean = optim_out$par[1], sd = optim_out$par[2])
    dist_data <- pnorm(seq(-0.3, 0.3, length.out = 600), mean = optim_out$par[1], sd = optim_out$par[2])
    dens_data <- dnorm(seq(-0.3, 0.3, length.out = 600), mean = optim_out$par[1], sd = optim_out$par[2])
    theoretical_quantiles <- qstd(ppoints(n), mean = optim_out$par[1], sd = optim_out$par[2])
  }
  
  r_squared <- cor(sort(fit), sort(x))
  r_squared_round <- round(r_squared, 3)
  
  aic = 2 * optim_out$value + 2 * 4
  bic = 2 * optim_out$value + log(n) * 4
    
  sample_mean <- mean(x)
  
  if(distribution == "sstd") {
    qq_subtitle <- paste0(
      "m=", round(optim_out$par[1], 4), 
      ", s=", round(optim_out$par[2], 4),
      ", nu=", round(optim_out$par[3], 4),
      ", xi=", round(optim_out$par[4], 4),
      ", R^2=", r_squared_round
    )
  } else if(distribution == "std") {
    qq_subtitle <- paste0(
      "m=", round(optim_out$par[1], 4), 
      ", s=", round(optim_out$par[2], 4),
      ", nu=", round(optim_out$par[3], 4),
      ", R^2=", r_squared_round
    )
  } else {
    qq_subtitle <- paste0(
      "m=", round(optim_out$par[1], 4), 
      ", s=", round(optim_out$par[2], 4),
      ", R^2=", r_squared_round
    )
  }

  output <- list(
    # Plot with estimated skewed t-distribution
    # qq plot for t-distribution
    qqplot = function() {
      qqplot(
        x = theoretical_quantiles, 
        y = x, 
        main = paste0("QQ-plot, ", distribution),
        xlab = paste0(distribution, " quantiles"), 
        ylab = "log returns", 
        xlim = c(min(fit) - abs(min(fit))/10, max(fit) + abs(max(fit))/10),  
        ylim = c(min(x) - abs(min(x))/10, max(x) + abs(max(x))/10)
      )
      mtext(
        qq_subtitle, 
        side=3,  
        cex = 0.7, 
        adj=0
      )
      qqline(
        x, 
        distribution = function(p) {
          if(distribution == "sstd") {
            qsstd(
              p, optim_out$par[1], optim_out$par[2], optim_out$par[3], optim_out$par[4]
            )
          } else if (distribution == "std") {
            qstd(
              p, optim_out$par[1], optim_out$par[2], optim_out$par[3]
            )
          } else {
            qnorm(p, optim_out$par[1], optim_out$par[2])
          }
        }, 
        datax = FALSE, 
        col="black"
      )
      abline(0, 1, col = "red")
      legend(
        "bottomright", 
        legend = c("data trendline", "45 degree line"), 
        col = c("black", "red"), 
        lty = c(1, 1)
      )
    },
    fit_plot = function() {
      plot(
        sort(x), 
        col="blue", 
        ylab = "log returns", 
        main = "Data vs fit",
        ylim = c(min(c(x, fit)), max(c(x, fit)))
      )
      points(
        sort(fit), 
        col="red"
        )
      legend(
        "bottomright", 
        legend = c("data", "fit"), 
        col = c("blue", "red"), 
        pch = c(1, 1)
      )
    },
    quantile_plot = function() {
      plot(
        x = seq(0.005, 0.995, length.out = 600), 
        sort(quantile_data), 
        pch = 16, 
        cex = 0.3, 
        main = paste0("Estimated ", distribution, " distribution quantiles"),
        xlab = "probability", ylab = "log-returns")
      abline(a = 0, b = 0, col = "gray", lty = 2)
    },
    dist_plot = function() {
      plot(
        x = seq(-0.3, 0.3, length.out = 600), 
        sort(dist_data), 
        pch = 16, 
        cex = 0.3, 
        main = paste0("Estimated ", distribution, " distribution CDF"),
        xlab = "log-returns", ylab = "probability")
      abline(v = c(min(x), max(x)), col = c("red", "green"))
      abline(
        h = if(distribution == "sstd") {
          c(
            psstd(min(x), optim_out$par[1], optim_out$par[2], optim_out$par[3], optim_out$par[4]), 
            psstd(max(x), optim_out$par[1], optim_out$par[2], optim_out$par[3], optim_out$par[4])
          )
        } else if (distribution == "std") {
          c(
            pstd(min(x), optim_out$par[1], optim_out$par[2], optim_out$par[3]), 
            pstd(max(x), optim_out$par[1], optim_out$par[2], optim_out$par[3])
          )
        } else {
          c(
            pnorm(min(x), optim_out$par[1], optim_out$par[2]), 
            pnorm(max(x), optim_out$par[1], optim_out$par[2])
          )
        }, 
        col = c("red", "green")
      )
    },
    dens_plot = function() {
      plot(
        x = seq(-0.3, 0.3, length.out = 600), 
        y = dens_data, 
        cex = 0.3, 
        pch = 16, 
        xlab = "log-return", 
        ylab = "likelihood", 
        main = paste0("Estimated ", distribution, " distribution PDF"),
        sub = paste0("m = ", round(optim_out$par[1], 3), ", sample mean = ", round(sample_mean, 3)))
      abline(v = optim_out$par[1], col = "red")
      abline(v = sample_mean, lty = 2, col = "blue")
      legend(
        "topleft", 
        legend = c("density", "m", "sample mean"), 
        col = c("black", "red", "blue"), 
        lty = c(1, 1, 2)
      )
    },
    data = x,
    fit = fit,
    dist_params = optim_out$par,
    dist_data = dist_data,
    dens_data = dens_data,
    quantile_data = quantile_data,
    theoretical_quantiles = theoretical_quantiles,
    r_squared = r_squared,
    aic = aic,
    bic = bic
  )
  
  output
}

fit_distribution_params <- function(x) {
  suppressWarnings(fit_distribution(x)$dist_params)
}
```

```{r}
#' @param x A numeric vector, matrix, or data frame of time series data.
#' @param block_length Length of individual blocks (ignored for MSAR).
#' @param block_type Type of block sampling: "moving", "stationary", "hmm", 
#'   "msar", or "wild".
#' @param func A function to apply to each bootstrap sample (e.g., mean).
#' @param apply_func_to Apply function to "cols" (columns individually) or "df" (whole data.frame).
#' @param num_blocks Number of blocks per bootstrap sample.
#' @param num_boots Number of bootstrap replicates.
#' @param p_method Method to estimate average block probability p: "1/n",
#'   "plugin", or "cross validation".
#' @param p Optional value for p. If NULL, uses p_method to estimate.
#' @param overlap Whether to allow overlapping blocks (for moving/stationary).
#' @param ar_order Order of AR model used in MSAR block segmentation.
#' @param nstates Number of states in the MSAR model.
#' @param model_func Model function used when p_method = "cross validation".
#' @param score_func Scoring function used when p_method = "cross validation".
vanilla_boot <- bootstrap(
  returns, 
  n_length = 252,
  block_length = 1, 
  block_type = "stationary",
  func = fit_distribution_params, 
  apply_func_to = "cols",
  num_blocks = 252, 
  num_boots = 100, 
  p_method = "1/n",
  p = NULL, 
  overlap = TRUE, 
  ar_order = 1, 
  nstates = 2,
  model_func = NULL, 
  score_func = NULL
)



```





