---
title: "Starter System"
author: "mhv"
date: "6/29/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(timeDate)
```


# Starter system


## Get data: Single instrument, single product
### Price data
30 years of simulated price data
```{r}
holidays <- holidayNYSE()
days <- as.timeDate(seq(from = as.Date("2000-01-01"), to = as.Date("2021-12-31"), by = "day"))
bizdays <- days[isBizday(days, holidays = holidays, wday = 1:5)]

data <- data.frame(date = bizdays)
names(data)[1] <- "date"

set.seed(3630)
data$returns <- c(NA, rnorm(nrow(data) - 1, 0, 0.05/sqrt(252)))
#data$price <- cumprod(c(100, data$returns[2:length(data$returns)] + 1)) ## Initial price is 100
data$price <- prices_from_returns(data$returns, 100)
```

```{r}
plot(data$date, data$price, pch = 16, cex = 0.2, xlab = "date", ylab = "price")
```

## Configuration
```{r}
data$capital[1] <- 1000000
mac <- c(16, 64) ## moving average crossover
sr <- 0.24 ## Sharpe Ratio
num_trades <- 5.4 ## Expected number of trades per year
risk_target <- 0.12

stop_loss_fraction <- 0.5
risky_capital_pct_ <- risky_capital_pct(risk_target, stop_loss_fraction)
#stop_loss_gap_ <- 

cat("Capital:", capital,"\n")
cat("Moving average crossover:", mac[1], ",", mac[2],"\n")
cat("Sharpe Ratio:", sr,"\n")
cat("Expected number of trades per year:", num_trades,"\n")
cat("Risk target:", risk_target,"\n")
cat("Instrument risk in %:", instrument_risk_,"\n")
cat("Notional exposure:", notional_exposure_,"\n")
cat("\n")
cat("Stop loss fraction:", 0.5)
```

## Moving Average

MA16
```{r}
data$ma16 <- moving_average(data$price, 16)
```

MA64
```{r}
data$ma64 <- moving_average(data$price, 64)
```

MA256
```{r}
data$ma256 <- moving_average(data$price, 256)
```

```{r}
plot(data$date, data$price, pch = 16, cex = 0.2, xlab = "date", ylab = "price")
points(data$date, data$ma16, pch = 16, cex = 0.2, col = "green")
points(data$date, data$ma64, pch = 16, cex = 0.2, col = "red")
points(data$date, data$ma256, pch = 16, cex = 0.2, col = "blue")
```


## Moving Average Crossover

```{r}
data$mac <- moving_average_crossover(data$ma64, data$ma256)
```


Gray dot = 1 indicates uptrend.  
Gray dot = 0 indicates downtrend.
```{r}
plot(data$date, data$price, pch = 16, cex = 0.2, xlab = "date", ylab = "price", ylim = c(min(data$price), max(data$price)))
points(data$date, data$ma64, pch = 16, cex = 0.2, col = "red")
points(data$date, data$ma256, pch = 16, cex = 0.2, col = "blue")
points(data$date, data$mac * (max(data$price) - min(data$price)) + min(data$price), pch = 16, cex = 0.2, col = "gray")
```
# Returns
```{r}
plot(data$date[2:length(data$date)], data$returns[2:length(data$date)], type = "l", lwd = 0.2, xlab = "date", ylab = "return")
```

# Trade simulation 1
```{r}
data$instr_risk <- rep(NA, nrow(data))
data$notional_exposure <- rep(NA, nrow(data))
data$position_size <- rep(NA, nrow(data))
for(t in 26:(nrow(data))) {
  data$instr_risk[t] <- instr_risk(data$price[(t-25):t]) ## Annualised standard deviation of last 25 returns
  data$notional_exposure[t] <- notional_exposure(risk_target, prices$capital[t], data$instr_risk[t])
  data$position_size[t] <- data$notional_exposure[t] / data$price[t]
}
```


```{r}
plot(data$date[26:length(data$date)], data$instr_risk[26:length(data$date)], type = "l", lwd = 0.2, xlab = "date", ylab = "instrument risk")
```

```{r}
plot(data$date[26:length(data$date)], data$notional_exposure[26:length(data$date)], type = "l", lwd = 0.2, xlab = "date", ylab = "notional exposure")
```

```{r}
plot(data$date[26:length(data$date)], data$position_size[26:length(data$date)], type = "l", lwd = 0.2, xlab = "date", ylab = "position size")
```

# Trade simulation 2

Generate time series `prices` of simulated prices.  
Trades simulation iteratively builds a data frame `trades_data` with `prices` and previous `trades_data` as input.

```{r}
volatility = 0.12 ## I chose same vol as target_risk

holidays <- holidayNYSE()
days <- as.timeDate(seq(from = as.Date("2000-01-01"), to = as.Date("2021-12-31"), by = "day"))
bizdays <- days[isBizday(days, holidays = holidays, wday = 1:5)]

prices <- data.frame(date = bizdays)
names(prices)[1] <- "date"

#set.seed(3630)
set.seed(1729)
prices$returns <- c(NA, rnorm(nrow(prices) - 1, 0, volatility/sqrt(252)))

prices$price <- prices_from_returns(prices$returns, 100)
```

Replace prices with neat integer series
```{r}
gen_num <- function(max_num) {
  num = NA
  flag_up = 1
  for(t in 1:max_num){
    flag_down = 1
    
    if(flag_up == 1) {
      num[t] <- ((t-1)%%200) + 1
      flag_down = 0
      if(num[t] == 200) {
        flag_up = 0
      }
    }
    
    if(flag_down == 1) {
      num[t] <- 200 - (t-1)%%200 + 1
      flag_up = 0
      if(num[t] == 2) {
        flag_up = 1
      }
    }
  }
  num
}
prices$price <- gen_num(nrow(prices))
```


Run the simulation...
```{r}
output <- trade_sim(prices, init_capital = 1000, n_fast = 32, n_slow = 128, risk_target = 0.12, stop_loss_fraction = 0.5)

trades <- output[[1]]
accounts <- output[[2]]
```


```{r}
trades
```

```{r}
accounts
```


```{r}
plot(trades$date, trades$price, pch = 16, cex = 0.3, xlab = "date", ylab = "dollars")
points(trades$date, trades$ma_fast, pch = 16, cex = 0.25, col = "red")
points(trades$date, trades$ma_slow, pch = 16, cex = 0.2, col = "green")
points(trades$date, (trades$mac + 1) * 0.5 * (max(trades$price) - min(trades$price)) + min(trades$price), pch = 16, cex = 0.2, col = "gray")
title('MA(32, 128)')
legend(
  x = as.numeric(trades$date[round(length(trades$date) * 0.8)]), y = max(trades$price) * 0.9,  
  legend=c('price', 'ma_short', 'ma_long'), 
  col=c('black','red','green'), 
  lty=c(1,1,1), cex=0.6
)
```

```{r}
plot(trades$date, accounts$pnl, pch = 16, cex = 0.4, xlab = "date", ylab = "dollars", ylim = c(min(c(accounts$pnl, accounts$capital, accounts$cash)), max(c(accounts$pnl, accounts$capital, accounts$cash))))
points(trades$date, accounts$capital, pch = 16, cex = 0.4, col = "red")
points(trades$date, accounts$cash, pch = 16, cex = 0.15, col = "green")
legend(
  x = as.numeric(trades$date[round(length(trades$date) * 0.8)]), y = max(c(accounts$pnl, accounts$capital, accounts$cash)) * 0.9, 
  legend=c('pnl', 'capital', 'cash'), 
  col=c('black','red','green'), 
  lty=c(1,1,1), cex=0.6
)
```

(Lines here are not dashed. The gaps are weekends.)
```{r}
plot(trades$date[1:500], accounts$pnl[1:500], pch = 16, cex = 0.2, xlab = "date", ylab = "dollars", ylim = c(min(accounts$pnl[1:500], accounts$capital[1:500], accounts$cash[1:500]), max(accounts$pnl[1:500], accounts$capital[1:500], accounts$cash[1:500])))
points(trades$date[1:500], accounts$capital[1:500], pch = 16, cex = 0.3, col = "red")
points(trades$date[1:500], accounts$cash[1:500], pch = 16, cex = 0.15, col = "green")
title('pnl, capital, cash (first 500 days)')
legend(
  x = as.numeric(trades$date[400]), y = max(accounts$pnl[1:500], accounts$capital[1:500], accounts$cash[1:500]) * 0.9,  
  legend=c('pnl', 'capital', 'cash'), 
  col=c('black','red','green'), 
  lty=c(1,1,1), cex=0.6
)
```
```{r}
plot(trades$date, trades$notional_exposure, pch = 16, cex = 0.2, xlab = "date", ylab = "dollars")
title('notional exposure')
```

