---
title: "Portfolio selection - development"
author: Martin Hoshi Vognsen
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
   - \usepackage[fontsize=8pt]{scrextend}
mainfont: SourceSansPro
output: 
  html_document:
    toc: true
    toc_depth: 3
    keep_md: yes
  # pdf_document:
  #   toc: true
  #   toc_depth: 3
#fontsize: 10pt # for pdf. Limited to 10pt, 11pt and 12pt. Else use scrextend.
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
---


# Portfolio selection


```{r}
f_sharpe_ratio <- function(mu_p, mu_free, sigma) {
  (mu_p - mu_free) / sigma
}
```


## Simple mean-variance optimization w/o constraints

See "Lecture Notes for Finance 1 (and More), May 12, 2015", by Lando/Nielsen/Poulsen, pp. 154ff.  
The formula for the tangency portfolio weights does not appear in the notes but can easily be derived by setting $\mu_{\text{pf}}^{\text{e}}$ to the expression for $\mu_{\text{tan}}^{\text{e}}$ from the lecture notes p. 162.

Weights of the tangency portfolio:
$$\mathbf{w}_{\text{pf}}^{\text{e}} = \dfrac{1}{(\mathbf{\mu}^{\text{e}})^{\text{T}} \mathbf{\Sigma^{-1}} \mathbf{1}} \mathbf{\Sigma^{-1}} \mathbf{\mu}^{\text{e}}$$
Expected rate of return of the tangency portfolio:
$$\mathbf{\mu}_{\text{tan}}^{\text{e}} = \dfrac{(\mathbf{\mu}^{\text{e}})^{\text{T}} \mathbf{\Sigma^{-1}} \mathbf{\mu}^{\text{e}}}{\mathbf{1}^{\text{T}}  \mathbf{\Sigma^{-1}} \mathbf{\mu}^{\text{e}}}$$

Variance of rate of return of the tangency portfolio:
$$\sigma_{\text{tan}}^2 = \frac{(\mathbf{\mu}^{\text{e}})^{\text{T}} \mathbf{\Sigma^{-1}} \mathbf{\mu}^{\text{e}}}{(\mathbf{1}^{\text{T}}  \mathbf{\Sigma^{-1}} \mathbf{\mu}^{\text{e}})^2} = \dfrac{\mathbf{\mu}_{\text{tan}}^{\text{e}}}{\mathbf{1}^{\text{T}}  \mathbf{\Sigma^{-1}} \mathbf{\mu}^{\text{e}}}$$

### Tangency pf w/o constraints
```{r}
tangency_pf_simple <- function(
    mu_excess, ## mean excess rate of return  
    Sigma_inv ## Inverse covariance matrix, i.e. solve(Sigma)
  ) {
  
  ones <- rep(1, length(mu_excess))
  
  weights <- (Sigma_inv %*% mu_excess) / ((t(mu_excess) %*% Sigma_inv) %*% ones)
  
  denom <- (t(ones) %*% Sigma_inv) %*% mu_excess
  mu_excess_tan <- ((t(mu_excess) %*% Sigma_inv) %*% mu_excess) / denom
  
  sigma_excess_tan <- mu_tan / denom
  
  list(
    weights = weights,
    mu_excess_tan = mu_excess_tan, ## Mean excess return rate of tangency pf
    sigma_excess_tan = sigma_tan_excess ## Sigma of return rate of tangency pf
  )
}
```

### Tangency pf with constraints
```{r}

```



