---
title: "Portfolio selection - exploratory"
author: Martin Hoshi Vognsen
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
   - \usepackage[fontsize=8pt]{scrextend}
mainfont: SourceSansPro
output: 
  html_document:
    toc: true
    toc_depth: 3
    keep_md: yes
  # pdf_document:
  #   toc: true
  #   toc_depth: 3
#fontsize: 10pt # for pdf. Limited to 10pt, 11pt and 12pt. Else use scrextend.
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
---

```{r chunk1, setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate) 
#py_install("tslearn")
```


## Getting data

Here are a few examples of how to get some free data.  
For the present analysis we will download a data set from Kaggle, but first a few other examples...

### Yahoo Finance

Use screener to find tickers:  
https://finance.yahoo.com/research-hub/screener/ 

#### quantmod

```{r chunk2, results='hide', message=FALSE}
library(tidyverse)
library(quantmod)
library(ggplot2)
```
 
Simplest call. "AAPL" object will be created

```{r chunk3, eval=FALSE}
getSymbols("AAPL")
AAPL %>% head
```

```{r chunk4, eval=FALSE}
getSymbols(c("AAPL", "IBM"))
IBM %>% head
AAPL %>% head
```

Example without auto assign
```{r chunk5, eval=FALSE}
d <- getSymbols("AAPL", auto.assign = F)
d %>% head
```
 
Specifying from/to dates
```{r chunk6, eval=FALSE} 
d <- getSymbols("AAPL", auto.assign = F, from = "2020-01-01", to = "2020-06-01")
d %>% head 
```
 
Specifying return class
```{r chunk7, eval=FALSE}
d <- getSymbols("AAPL", auto.assign = F, from = "2020-01-01", to = "2020-06-01", return.class='data.frame')
d %>% head  
```

Plot log-returns
```{r}
quantmod::getSymbols( "^GSPC", from="2004-01-01" )
gspcRets = diff( log( quantmod::Cl( GSPC ) ) )
xxx = as.numeric(gspcRets)
```



#### yahoofinancer

```{r chunk8, eval=FALSE, message=FALSE}
library(yahoofinancer)
```

https://yahoofinancer.rsquaredacademy.com/ 

https://cloud.r-project.org/web/packages/yahoofinancer/yahoofinancer.pdf

```{r chunk9, eval=FALSE}
aapl <- Ticker$new('aapl')
```

```{r chunk10, eval=FALSE}
names(aapl)
```


```{r chunk11, eval=FALSE}
aapl$regular_market_price
```

```{r chunk12, eval=FALSE}
names(aapl$technical_insights)
```



```{r chunk13, eval=FALSE}
aapl$fifty_two_week_high
```

```{r chunk14, eval=FALSE}
aapl$previous_close
```


```{r chunk15, eval=FALSE}
nifty_50 <- Index$new('^NSEI')
head(nifty_50$get_history(start = '2024-01-20', interval = '1d'))
```

```{r chunk16, eval=FALSE}
head(currency_converter('GBP', 'USD', '2024-01-20', '2024-01-30'))
```

```{r chunk17, eval=FALSE}
head(aapl$get_history(start = '2024-10-20', interval = '1d'))
```

```{r chunk18, eval=FALSE}
get_trending(country = "US", count = 10)
```


```{r chunk19, eval=FALSE}
wheat <- Ticker$new('CWDH25.CBT')
wheat$fifty_two_week_high
```



### Alpha Vantage

See Alpha Vantage SPI documentation:  
https://www.alphavantage.co/documentation/

NOTE  
25 requests per day limit  
https://www.alphavantage.co/premium/

#### Without `alphavantager`:

```{r chunk20, eval=FALSE, message=FALSE}
library(tidyverse)
library(jsonlite)
library(httr)
 
symbol = "AMZN"
av_key = ""
 
url <- str_c("https://www.alphavantage.co/query?function=TIME_SERIES_DAILY&symbol=", symbol ,"&apikey=", av_key, "&datatype=csv")
 
d <- read_csv(url)
 
d %>% head
```

#### With `alphavantager` package

SETUP API KEY
```{r chunk21, eval=FALSE}
av_api_key(av_key)
```


TIME SERIES INTRADAY
```{r chunk22, eval=FALSE}
av_get("MSFT", av_fun = "TIME_SERIES_INTRADAY", interval = "5min", outputsize = "full")
```

Get market cap of Amazon
```{r chunk23, eval=FALSE}
d <- av_get(symbol = "AMZN", av_fun = "OVERVIEW")
d$value[which(d$rank_group == "MarketCapitalization")]
```

## Top market cap tickers

Got 480 top market cap tickers from  
https://www.marketwatch.com/tools/screener/stock?exchange=all&skip=0&orderbyfield=Symbol&direction=asc&visiblecolumns=Symbol&marketcapmin=25000.0&marketcapmax=10000000.0

NOTE  
Cap data is available from  
- https://www.nasdaq.com/market-activity/stocks/screener?page=1&rows_per_page=25  
	- See here: `/Users/mhvpbp13/Library/Mobile Documents/com~apple~CloudDocs/Data research/_data/ticker symbols`

Then manually copy-pasted tickers into csv. (Could not figure out how to scrape.)

```{r chunk24}
top_mcap_tickers <- as.vector(read.csv("/Users/mhvpbp13/Library/Mobile\ Documents/com~apple~CloudDocs/Data\ research/aktieanalyser/top_marketcap.csv"))[[1]]
```

```{r chunk25}
top_mcap_tickers
```


Many data sets are not available, because the data only runs until 2018.

```{r chunk26, warning=FALSE, results='hide'}
top_mcap_data <- lapply(top_mcap_tickers, function(ticker) {
  try(
    df <- read.csv(
      paste0(
        "/Users/mhvpbp13/Library/Mobile\ Documents/com~apple~CloudDocs/Data\ research/_data/kaggle/full_history/", ticker, ".csv"
      )
    )
  )
})

names(top_mcap_data) <- top_mcap_tickers
```


How many data sets succeded?

```{r chunk27}
test_ticker_data <- as.vector(t(as.data.frame(lapply(top_mcap_data, function(x) {
    is.data.frame(x) && nrow(x) != 0
  }))))
sum(test_ticker_data)
```

Make list of only successful tickers
```{r chunk28}
top_mcap_data_valid <- top_mcap_data[which(test_ticker_data)]
```


Make a list of price series for all valid tickers (close prices).  
Data sets are in order newest at the top, so read in reverse order.  
```{r chunk29}
top_mcap_prices_list <- lapply(top_mcap_data_valid, function(data) {
    zoo::read.zoo(data.frame(
      date = rev(data$date),
      price = rev(data$close)
    ))
  }
)
```

Combine the data frames of prices.  
Merging will put the each price in the the correct (time) row, because we
converted each data frame to a zoo object above.
```{r chunk30}
top_mcap_prices_df <- do.call(merge, top_mcap_prices_list)
```

Make a list of log-returs for all valid tickers (close prices)
```{r chunk31}
log_returns_from_prices <- function(prices) {
  N <- length(prices)
  log(utils::tail(prices, N - 1) / utils::head(prices, N - 1))
}

top_mcap_returns_list <- lapply(top_mcap_data_valid, function(data) {
    zoo::read.zoo(data.frame(
      date = rev(data$date)[-1],
      logreturn = log_returns_from_prices(rev(data$close))
    ))
  }
)
```


Combine the data frames (actually zoo matrices) of logreturns.  
Merging will put the each logreturn in the the correct (time) row, because we
converted each data frame to a zoo object above.
```{r chunk32}
top_mcap_returns_df <- do.call(merge, top_mcap_returns_list)
```

NOTE  
To turn a zoo matrix into a data frame with a "data" column, we could extract the matrix containing core data of the zoo object with `coredata()` and the time vector with `index()`.



We want all series to start at the first date of the shortest series.  
Which elements in each series are not NA? Pick the first.
```{r chunk33}
first_dates <- lapply(top_mcap_returns_df, function(x) {which(!is.na(x))[1]})
```

Which of all the _first dates_ is the latest one?
```{r chunk34}
first_date <- first_dates[which(first_dates == max(unlist(first_dates)))]
first_date
```

Let's start the data there...
```{r chunk35}
top_mcap_returns_df_short <- top_mcap_returns_df[unlist(first_date):(nrow(top_mcap_returns_df))]
```

Only two rows!  
Let's omit LIN...
```{r chunk36}
first_dates <- lapply(top_mcap_returns_df[, -which(names(top_mcap_returns_df) == "LIN")], function(x) {which(!is.na(x))[1]})
```

Which of all the _first dates_ is NOW the latest one?
```{r chunk37}
first_date <- first_dates[which(first_dates == max(unlist(first_dates)))]
first_date
```

Let's omit PDD as well...
Repeating this process some times gives us...
```{r chunk38}
omitted <- c(
  which(names(top_mcap_returns_df) == "LIN"),
  which(names(top_mcap_returns_df) == "PDD"),
  which(names(top_mcap_returns_df) == "KDP"),
  which(names(top_mcap_returns_df) == "SPOT"),
  which(names(top_mcap_returns_df) == "ZS"),
  which(names(top_mcap_returns_df) == "EQNR"),
  which(names(top_mcap_returns_df) == "MUFG"),
  which(names(top_mcap_returns_df) == "SE"),
  which(names(top_mcap_returns_df) == "VICI"),
  which(names(top_mcap_returns_df) == "ARGX"),
  which(names(top_mcap_returns_df) == "CVNA"),
  which(names(top_mcap_returns_df) == "TTD"),
  which(names(top_mcap_returns_df) == "FTV"),
  which(names(top_mcap_returns_df) == "TEAM"),
  which(names(top_mcap_returns_df) == "RACE"),
  which(names(top_mcap_returns_df) == "HPE"),
  which(names(top_mcap_returns_df) == "KHC"),
  which(names(top_mcap_returns_df) == "PYPL"),
  which(names(top_mcap_returns_df) == "AXON"),
  which(names(top_mcap_returns_df) == "SHOP"),
  which(names(top_mcap_returns_df) == "GDDY"),
  which(names(top_mcap_returns_df) == "QSR"),
  which(names(top_mcap_returns_df) == "KEYS"),
  which(names(top_mcap_returns_df) == "HUBS"),
  which(names(top_mcap_returns_df) == "BABA"),
  which(names(top_mcap_returns_df) == "SYF"),
  which(names(top_mcap_returns_df) == "FWONK"),
  which(names(top_mcap_returns_df) == "ANET"),
  which(names(top_mcap_returns_df) == "JD"),
  which(names(top_mcap_returns_df) == "ARES"),
  which(names(top_mcap_returns_df) == "HLT"),
  which(names(top_mcap_returns_df) == "VEEV"),
  which(names(top_mcap_returns_df) == "FI"),
  which(names(top_mcap_returns_df) == "CDW"),
  which(names(top_mcap_returns_df) == "IQV"),
  which(names(top_mcap_returns_df) == "ZTS"),
  which(names(top_mcap_returns_df) == "FWONA"),
  which(names(top_mcap_returns_df) == "ABBV"),
  which(names(top_mcap_returns_df) == "MPLX"),
  which(names(top_mcap_returns_df) == "FANG"),
  which(names(top_mcap_returns_df) == "WDAY"),
  which(names(top_mcap_returns_df) == "PANW"),
  which(names(top_mcap_returns_df) == "NOW"),
  which(names(top_mcap_returns_df) == "PSX"),
  which(names(top_mcap_returns_df) == "XYL"),
  which(names(top_mcap_returns_df) == "MPC"),
  which(names(top_mcap_returns_df) == "APO"),
  which(names(top_mcap_returns_df) == "HCA"),
  which(names(top_mcap_returns_df) == "KMI"),
  which(names(top_mcap_returns_df) == "TRGP"),
  which(names(top_mcap_returns_df) == "GM"),
  which(names(top_mcap_returns_df) == "LPLA"),
  which(names(top_mcap_returns_df) == "COR"),
  which(names(top_mcap_returns_df) == "NXPI"),
  which(names(top_mcap_returns_df) == "KKR"),
  which(names(top_mcap_returns_df) == "ET"),
  which(names(top_mcap_returns_df) == "UBS"),
  which(names(top_mcap_returns_df) == "VST")
  
)
        
first_dates <- lapply(top_mcap_returns_df[ , -omitted], function(x) {which(!is.na(x))[1]})

first_date <- first_dates[which(first_dates == max(unlist(first_dates)))]
first_date
```

Also removing UBS and VST, which look weird.

Ok, this gives us a bit more than 8 years of data...
```{r chunk39}
top_mcap_returns_df_short <- top_mcap_returns_df[unlist(first_date):(nrow(top_mcap_returns_df)), -omitted]
top_mcap_prices_df_short <- top_mcap_prices_df[(unlist(first_date) - 1):(nrow(top_mcap_prices_df)), -omitted]

cat(paste0("Number of rows: ", nrow(top_mcap_returns_df_short), "\n"))
cat(paste0("Number of assets: ", ncol(top_mcap_returns_df_short)))
```

How many NAs?
```{r chunk40}
sum(is.na(top_mcap_returns_df_short))
```

Ok then.

Interpolate NAs with zoo...  
```{r chunk41}
# replace trailing NA with previous value
replace_trailing_NAs <- function(x) {
  for(i in seq_along(x)) {
    if(is.na(x[i])) {x[i] <- x[i - 1]}
  }
  x
}

top_mcap_returns_df_short_filled <- top_mcap_returns_df_short

for(i in 1:ncol(top_mcap_returns_df_short)) {
  col <- top_mcap_returns_df_short[, i]
  if(sum(is.na(col)) > 0) {
    top_mcap_returns_df_short_filled[, i] <- zoo::na.approx(col, na.rm = FALSE)
    
    # cat(sum(is.na(col)), "\n")
    # cat(sum(is.na(top_mcap_returns_df_short_filled[, i])), "\n")

    if(sum(is.na(col)) == sum(is.na(top_mcap_returns_df_short_filled[, i]))) {
      # We should have removed all leading NAs, so any remaining NAs must be trailing...
      top_mcap_returns_df_short_filled[, i] <- replace_trailing_NAs(top_mcap_returns_df_short_filled[, i])
    }
  }
}
```

```{r chunk42}
sum(is.na(top_mcap_returns_df_short_filled))
```





## Clusters vs sectors

Now, let's try k-Sharp...


See 

+ https://www.rdocumentation.org/packages/dtwclust/versions/3.1.1/topics/dtwclust  
+ https://journal.r-project.org/archive/2019/RJ-2019-023/RJ-2019-023.pdf  


This can not be installed... (Fails to compile dependencies)  
Maybe try Linux box...?  
```{r chunk43, eval=FALSE, message=FALSE}
#library(dtwclust)
```


```{r chunk44, eval=FALSE}
# tsclust(top_mcap_returns_df[ ,-1], k = 4L, seed = 8L,
# distance = "sbd", centroid = "shape")
```



#### Sectors

We would like to get the sector for each ticker symbol.

We could try getting the data with `alphavantager` package...

SETUP API KEY
```{r  chunk46, eval=FALSE}
av_api_key(av_key)
```


TIME SERIES INTRADAY
```{r  chunk47, eval=FALSE}
av_get("MSFT", av_fun = "TIME_SERIES_INTRADAY", interval = "5min", outputsize = "full")
```

Get market cap of Amazon
```{r  chunk48, eval=FALSE}
d <- av_get(symbol = "AMZN", av_fun = "OVERVIEW")
d$value[which(d$rank_group == "MarketCapitalization")]
```


!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!  
OBS!  
Alpha Vantage had max 25 requests per day.  
Cluster 1 has 45 symbols - so 45 requests!!  
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!  

```{r chunk49}
range_of_symbols_to_get <- 1:2
```

```{r chunk50, eval=FALSE}
symbols_1 <- names(top_mcap_prices_df_short)[which(py$mcap_clusters == 0)]
overviews <- lapply(symbols_1[range_of_symbols_to_get], function(x) {av_get(symbol = x, av_fun = "OVERVIEW")})
sectors <- lapply(overviews, function(x) {x$value[which(x$rank_group == "Sector")]})
```


Instead we can download the data for free as csv here:

https://www.nasdaq.com/market-activity/stocks/screener?page=1&rows_per_page=25 

```{r chunk51}
nasdaq_table <- read.csv("/Users/mhvpbp13/Library/Mobile\ Documents/com~apple~CloudDocs/Data\ research/_data/ticker\ symbols/nasdaq_tickers.csv")
nyse_table <- read.csv("/Users/mhvpbp13/Library/Mobile\ Documents/com~apple~CloudDocs/Data\ research/_data/ticker\ symbols/nyse_tickers.csv")
#ticker_table <- rbind(nasdaq_table[, c("Symbol", "Sector")], nyse_table[, c("Symbol", "Sector")])
ticker_table <- rbind(nasdaq_table, nyse_table)
```



### Python in R with Reticulate

See https://rstudio.github.io/cheatsheets/html/reticulate.html


```{r}
num_clusters <- 10L
```


Note that runtime for k-Shape is linear in $n$, the number of time series, and linear in $k$, the number of clusters, but increases as $m^3$ for $m$ elements in each data set. The algorithm gets slow for $m > n$, 

```{python chunk54, warning=FALSE}
import tslearn.clustering as clust
```

For some reason it seems we have to transpose...
```{r chunk55}
top_mcap_returns_df_short_filled_matrix <- as.matrix(t(top_mcap_returns_df_short_filled))
```


Generate k-Shape clusters
```{python chunk56}
mcap_clusters_model = clust.KShape(n_clusters=r.num_clusters, max_iter=100, n_init=1, random_state=0)

mcap_clusters = mcap_clusters_model.fit_predict(r.top_mcap_returns_df_short_filled_matrix)
```
```{r chunk52, results='asis'}
## Cluster numbers start with 0.
## For presentation starting with 1 may make more sense.
## So if the input cluster is 0, the printed cluster number will be 1
group_by_sectors <- function(cluster_num, verbose = FALSE) {
  symbols <- names(top_mcap_prices_df_short)[which(py$mcap_clusters == cluster_num)]
  sectors <- ticker_table[which(ticker_table$Symbol %in% symbols), ]
  sectors_list <- lapply(unique(sectors$Sector), function(x) {
      list(sectors$Symbol[which(sectors$Sector == x)])
    }
  )
  names(sectors_list) <- unique(sectors$Sector)
  
  cat("\n")
  cat("====== Cluster ", cluster_num + 1, " ======\n")
  
  if(verbose == TRUE) {
    for(i in seq_along(sectors_list)) {
      cat("\n")
      cat(names(sectors_list[i]), ":\n")
      cat("\n")
      cat(unlist(sectors_list[i]), "\n")
      cat("\n")
    }
  }
  
  sectors_list
}
```


```{r chunk53, results='asis'}
for(i in 0:(num_clusters - 1)) {
  sector_lists <- group_by_sectors(cluster_num = i, verbose = TRUE)
}
```


```{python chunk57}
centroids = mcap_clusters_model.cluster_centers_
centroids.shape
```


```{r chunk58}
dim(py$centroids)
```

```{r chunk59}
centroids <- as.data.frame(t(py$centroids[,,1]))
```


It seems the Python function KShape doesn't provide the label of the centroids,
so we find the centroid ticker symbols by minimizing sum of square differences
between centroids (scaled by KShape) and return vectors for each cluster.
```{r chunk61}
find_centroid_ticker <- function(centroiddf) {
  
  centroid_symbols <- list()

  for(i in seq_along(centroiddf)) {
    
    id <- which(py$mcap_clusters == i - 1)
    cluster_symbols <-  names(top_mcap_returns_df_short)[id]
    cluster_returns <- scale(top_mcap_returns_df_short[, id])
    
    sum_sq_diffs <- numeric(ncol(cluster_returns))
    for(j in 1:ncol(cluster_returns)) {
      sq_diffs <- unlist((cluster_returns[, j] - centroiddf[, i])^2)
      sum_sq_diffs[[j]] <- sum(sq_diffs)
    }

    id <- which(sum_sq_diffs == min(sum_sq_diffs))
    centroid_symbols[i] <- if(length(cluster_symbols[id]) > 0) {
      cluster_symbols[id]} else {NA}
  }
 centroid_symbols
}

get_table_by_tickers <- function(tickers, tickertable) {
  tickertable1 <- tickertable[tickertable$Symbol %in% tickers, ]
  tickertable1[!is.na(tickertable1$Symbol), ]
}
```


Note: Why is one of them NA?
```{r chunk62}
centroid_symbols <- unlist(find_centroid_ticker(centroids))
centroid_symbols
```


```{r chunk63}
centroid_table <- get_table_by_tickers(centroid_symbols[!is.na(centroid_symbols)], ticker_table)
knitr::kable(centroid_table, digits = 3)
#centroid_table
```




```{r chunk64, results='hide', message=FALSE}
library(corrplot)
```


```{r chunk65}
centroid_correlations <- cor(centroids)
corrplot(centroid_correlations, method = "circle", tl.cex = 0.3)
```

```{r chunk66}
range(centroid_correlations[centroid_correlations != 1])
```



```{r chunk67}
centroid_symbols_ids <- which(names(top_mcap_prices_df_short) %in% centroid_symbols)
ggplot(
  aes(x = Index, y = Value, colour = Series), 
  data = fortify(scale(top_mcap_prices_df_short[, centroid_symbols_ids]), melt = TRUE)) + 
    geom_line(linewidth = 0.2) + 
    labs(title = "Scaled centroid prices", x ="Time", y = "Price") +
    theme(legend.position="none")
```

Plot weighted portfolio average of centroid stocks.  

```{r chunk68}
sds <- lapply(top_mcap_prices_df_short[, centroid_symbols_ids], function(x) {sd(x, na.rm = TRUE)})
#weights <- lapply(sds, function(x) {x/sum(unlist(sds))})
weights <- unlist(lapply(sds, function(x) {1/x}))
weights <- weights / sum(weights)

centroid_pf_prices <- as.matrix(top_mcap_prices_df_short[, centroid_symbols_ids]) %*% unname(unlist(weights))


centroid_portfolio <- data.frame(
  date = index(top_mcap_prices_df_short),
  price = centroid_pf_prices
)
  
  
ggplot(aes(x = date, y = price), data = centroid_portfolio) + 
    geom_line(linewidth = 0.2) + 
    labs(title = "Weighted avg. of centroid prices", x ="Time", y = "Price") +
    theme(legend.position="none")
```

## Cluster plots

```{r chunk45}
logreturn_sd <- unlist(lapply(top_mcap_returns_df_short_filled, sd))
logreturn_mean <- unlist(lapply(top_mcap_returns_df_short_filled, mean))
```

```{r chunk69}
topcapplotter <- function(cluster_num) {
  id <- which(py$mcap_clusters == cluster_num)
  max_sd <- max(logreturn_sd[id])
  
  tickers <- names(top_mcap_prices_df_short)[id]
  ticker_table <- get_table_by_tickers(tickers[!is.na(tickers)], ticker_table)
  
  if(length(id) > 0) {
    
    if(length(id) > 1) {
      correlations <- stats::cor(top_mcap_returns_df_short[, id])
      correlations_plot <- function() {corrplot(correlations, method = "circle", tl.cex = 0.3)}
    } else {
      correlations <- 1
      correlations_plot <- function() {NA}
    }
    
    price_plot <- ggplot(
        aes(x = Index, y = Value, colour = Series), 
        data = fortify(top_mcap_prices_df_short[, id], melt = TRUE)) +
      geom_line(linewidth = 0.2) +
      labs(title = paste0("Cluster ", cluster_num + 1), x ="Time", y = "Price") +
      theme(legend.position="none")
  
    index_plot <- ggplot(
         aes(x = Index, y = Value, colour = Series),
         data = fortify(100 * exp(cumsum(top_mcap_returns_df_short[, id])), melt = TRUE)) +
       geom_line(linewidth = 0.2) +
       labs(title = paste0("Cluster ", cluster_num + 1, " - indexed prices"), x ="Time", y = "Price index") +
       theme(legend.position="none")
  
    norm_ret_plot <- ggplot(
        aes(x = Index, y = Value, colour = Series),
        data = fortify(100 * exp(cumsum(scale(top_mcap_returns_df_short[, id]) * max_sd)), melt = TRUE)) +
      geom_line(linewidth = 0.2) +
      labs(title = paste0("Cluster ", cluster_num + 1, " - indexed prices from z-normalized logreturns"), x ="Time", y = "Price index") +
      theme(legend.position="none")
  
    norm_plot <- ggplot(
        aes(x = Index, y = Value, colour = Series),
        data = fortify(scale(100 * exp(cumsum(top_mcap_returns_df_short[, id]))), melt = TRUE)) +
      geom_line(linewidth = 0.2) +
      labs(title = paste0("Cluster ", cluster_num + 1, " - normalized indexed prices"), x ="Time", y = "Price index") +
      theme(legend.position="none")
  } else {
    tickers = NA
    id = NA
    price_plot = NA
    index_plot = NA
    norm_ret_plot = NA
    norm_plot = NA
    correlations = NA
    correlations_plot = function() {NA}
    ticker_table = NA
  }

  list(
    tickers = tickers,
    id = id,
    price_plot = price_plot,
    index_plot = index_plot,
    norm_ret_plot = norm_ret_plot,
    norm_plot = norm_plot,
    correlations = correlations,
    correlations_plot = correlations_plot,
    ticker_table = ticker_table
  )
}
```




```{r chunk69b, results='asis', eval=FALSE, include=FALSE, message=FALSE}
# Doesn't work, don't know why: Some chunks in the child doc work, but not all.  
# It works if I move the chunks from the child doc to the parent doc.  
library(here)
individual_cluster_plots <- lapply(
  1:num_clusters,
  function(i) {
    knitr::knit_child(
      here("misc", "notebooks", "Getting-data_clusterplots.Rmd"),
      envir = environment(), 
      quiet = TRUE
    )
  }
)

cat(unlist(individual_cluster_plots), sep = '\n')
```

```{r chunk70}
plots <- lapply(0:(num_clusters - 1), function(i) {topcapplotter(i)})
```

```{r chunk70a, eval=FALSE}
plots[[i]]$correlations_plot()
```

```{r chunk71, eval=FALSE}
cat("\n")
cat("Range of correlations: ", range(plots[[i]]$correlations[plots[[i]]$correlations != 1]), "\n")
cat("\n")
```


```{r chunk72, eval=FALSE}
cat(plots[[i]]$tickers)
```

```{r chunk73, eval=FALSE}
print(knitr::kable(plots[[i]]$ticker_table, digits = 3))
# plots$ticker_table
```


```{r chunk74, eval=FALSE}
print(plots[[i]]$price_plot)
```

```{r chunk75, eval=FALSE}
print(plots[[i]]$index_plot)
```

```{r chunk76, eval=FALSE}
print(plots[[i]]$norm_ret_plot)
```

```{r chunk77, eval=FALSE}
print(plots[[i]]$norm_plot)
```


```{r results='asis'}

for(i in 1:num_clusters) {
  cat("\n")
  cat("### Cluster ", i, "\n")
  cat("\n")
  if(length(plots[[i]]$tickers) > 0) {

    <<chunk70a>>
    cat("\n")
    <<chunk71>>
    cat("\n")
    <<chunk73>>
    cat("\n")
    <<chunk74>>
    cat("\n")
    <<chunk75>>
    cat("\n")
    <<chunk76>>
    cat("\n")
    <<chunk77>>
    cat("\n")
    
  } else {
    cat("\n")
    cat("Cluster ", i, " is empty.")
    cat("\n")
  }
}

```


## PCA

References:  

- Rupert: Statistics and Data Analysis for Financial Engineering
- https://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/112-pca-principal-component-analysis-essentials/#eigenvalues-variances
- https://ir.canterbury.ac.nz/server/api/core/bitstreams/a7b99f2d-5d9b-4212-a947-31c1d162d1c4/co

### Window length

Use KMO (Kaiser-Meyer-Olkin) statistic to test optimal window length.  
KMO is the ratio of correlations between stocks and those of the partial correlations.

```{r}
library(EFAtools) ## For KMO
```

2 is the smallest whole number of years that doesn't produce a singular correlation matrix.
```{r}
num_years <- 2
cor_matrix <- cor(top_mcap_returns_df_short[1:round(252 * num_years), ])
```

Find near perfect correlations, which make correlation matrix singular.
```{r}
perfect_correlations <- list()
for(i in 1:ncol(cor_matrix)) {
  perfect_correlations[[i]] <- which(cor_matrix[, i] > 0.99) != i
}
which(unlist(perfect_correlations))
```


Remove highly correlated stocks.  

```{r}
## TODO: Only remove one of each correlated pair.  
#symbols_to_remove <- names(unlist(perfect_correlations)[which(unlist(perfect_correlations))]) 

## Remove manually
symbols_to_remove <- "GOOG" 

cor_matrix_nonsingular <- cor_matrix[
  -which(rownames(cor_matrix) %in% (symbols_to_remove)), 
  -which(colnames(cor_matrix) %in% (symbols_to_remove))
]
```



Calculate KMO for different window lengths
```{r}
KMO(cor_matrix)
```

### Perform PCA
```{r}
pca_output <- prcomp(top_mcap_returns_df_short[1:504, ])
```

```{r}
summary(pca_output)
```

```{r}
library(FactoMineR)
```



Calculating PCA with z-normalization (centering and scaling).
```{r}
pca_output <- PCA(top_mcap_returns_df_short[1:504, ], scale.unit = TRUE, ncp = 10, graph = TRUE, axes = c(1,2))
```
```{r}
print(pca_output)
```

"cos2" is the quality of representation for variables on the factor map.  
"contrib" or "ctr" is the contributions of the variables.

```{r}
summary(pca_output)
```



### Scree plot

```{r}
library(factoextra)
```

The first PC captures half the total variation. The first PC is considered to capture systematic variation (market variation).  

The the first 10 PCs catches 60 percent of the total variation, and only 10 percent of the non-systematic variance, if the first PC represents the market variation.  

To capture 90 percent of the variation we would need 125 PCs.  

So the diversification provided by the first 10 PCs is not great.  

Following Kaiser's rule we would have to keep no more than 48 PCs (eigenvalue > 1).

```{r}
head(get_eigenvalue(pca_output), 125)
```


```{r}
fviz_eig(pca_output, addlabels = TRUE)
```

### Variables (individual stocks)

```{r}
library("corrplot")
```


```{r}
variables <- get_pca_var(pca_output)
```


Plot contribution of representation of the variables on factor map.  
Plotting the top 20 highest contributors to PC1.  

```{r}
top_pca_contrib_var <- head(variables$contrib[order(variables$contrib[, 1], decreasing=TRUE), ], 20)
# top_pca_variables <- lapply(variables, function(x) {
#   x[1:20, ]
# })
corrplot(top_pca_contrib_var, is.corr=FALSE)
```

Which stock contributes most to each of the first 10 PCs?
```{r}
top_contributing_stocks <- list()
for(i in 1:ncol(variables$contrib)) {
  top_contributing_stock_id <- which(variables$contrib[ ,i] == max(variables$contrib[,i]))
  top_contributing_stocks[[i]] <- variables$contrib[top_contributing_stock_id, i]
  names(top_contributing_stocks[[i]]) <- rownames(variables$contrib)[top_contributing_stock_id]
}
```

```{r}
head(top_contributing_stocks, 10)
```



Plot quality of representation of the variables on factor map.  
Plotting the top 20 highest contributors to PC1 and omitting PC1 in the plot, which otherwise dominates the plot.  
```{r}
top_pca_cor2_var <- head(variables$cos2[order(variables$contrib[, 1], decreasing=TRUE), ], 20)
# top_pca_variables <- lapply(variables, function(x) {
#   x[1:20, ]
# })
corrplot(top_pca_cor2_var[, -1], is.corr=FALSE)
```




Total cos2 of variables on dim. 1:
```{r}
fviz_cos2(pca_output, choice = "var", axes = 1) +
  theme(axis.text.x = element_blank())
```

Total contrib of variables on dim. 1.  
Red dashed is average.
```{r}
fviz_contrib(pca_output, choice = "var", axes = 1) +
  theme(axis.text.x = element_blank())
```


Total cos2 of variables on dim. 2:
```{r}
fviz_cos2(pca_output, choice = "var", axes = 2) +
  theme(axis.text.x = element_blank())
```

Total contrib of variables on dim. 2.  
```{r}
fviz_contrib(pca_output, choice = "var", axes = 2) +
  theme(axis.text.x = element_blank())
```


Total cos2 of variables on dim. 3:
```{r}
fviz_cos2(pca_output, choice = "var", axes = 3) +
  theme(axis.text.x = element_blank())
```

Total contrib of variables on dim. 3.  
```{r}
fviz_contrib(pca_output, choice = "var", axes = 3) +
  theme(axis.text.x = element_blank())
```


Total cos2 of variables on dim 4 thru 10.
```{r}
fviz_cos2(pca_output, choice = "var", axes = 4:10) +
  theme(axis.text.x = element_blank())
```

Total contrib of variables on dim. 4 thru 10.  
```{r}
fviz_contrib(pca_output, choice = "var", axes = 4:10) +
  theme(axis.text.x = element_blank())
```


We see that a lot of variables (stocks) contribute to PC1, but only a few higher PC numbers.


## Plot coordinates (eigenvectors)

First 10 components (dimensions):
```{r}
as.data.frame(pca_output$var$coord[, 1:10]) %>% 
  mutate(stock = factor(row.names(pca_output$var$coord))) %>% 
  gather(key = "component", value = "coord", -stock) %>% 
  ggplot(aes(x = stock, y = coord, group = component, colour = component)) +
    geom_line(linewidth = 0.2) +  
    geom_point(size = 0.5) +
    theme(axis.text.x = element_blank())
```

It is very clear to see, that all stockes - except a few - contribute mainly to $PC_1$, hence $PC_1$ represents market risk.

For which stocks are the coordinates for $PC_1$ smaller than for some other PC?

```{r}
strong_idio_risk <- list()
for(i in 2:10) {
  coords_i <- pca_output$var$coord[ ,i]
  strong_idio_risk_i <- coords_i[which(coords_i > pca_output$var$coord[ ,1])]
  strong_idio_risk_i
  if(length(strong_idio_risk_i) > 0) {
    strong_idio_risk[[i - 1]] <- strong_idio_risk_i
  } else {
    strong_idio_risk[[i - 1]] <- NA
  }
}
names(strong_idio_risk) <- paste0("component", 2:10)

strong_idio_risk
```

Let's look at the components in pairs.

First two components:
```{r}
as.data.frame(pca_output$var$coord[, 1:2]) %>% 
  mutate(stock = factor(row.names(pca_output$var$coord))) %>% 
  gather(key = "component", value = "coord", -stock) %>% 
  ggplot(aes(x = stock, y = coord, group = component, colour = component)) +
    geom_line(linewidth = 0.2) +  
    geom_point(size = 0.5) +
    theme(axis.text.x = element_blank())
```

Components 3 - 4:
```{r}
as.data.frame(pca_output$var$coord[, 3:4]) %>% 
  mutate(stock = factor(row.names(pca_output$var$coord))) %>% 
  gather(key = "component", value = "coord", -stock) %>% 
  ggplot(aes(x = stock, y = coord, group = component, colour = component)) +
    geom_line(linewidth = 0.2) +  
    geom_point(size = 0.5) +
    theme(axis.text.x = element_blank())
```

Components 5 - 6:
```{r}
as.data.frame(pca_output$var$coord[, 5:6]) %>% 
  mutate(stock = factor(row.names(pca_output$var$coord))) %>% 
  gather(key = "component", value = "coord", -stock) %>% 
  ggplot(aes(x = stock, y = coord, group = component, colour = component)) +
    geom_line(linewidth = 0.2) +  
    geom_point(size = 0.5) +
    theme(axis.text.x = element_blank())
```

Components 7 - 8:
```{r}
as.data.frame(pca_output$var$coord[, 7:8]) %>% 
  mutate(stock = factor(row.names(pca_output$var$coord))) %>% 
  gather(key = "component", value = "coord", -stock) %>% 
  ggplot(aes(x = stock, y = coord, group = component, colour = component)) +
    geom_line(linewidth = 0.2) +  
    geom_point(size = 0.5) +
    theme(axis.text.x = element_blank())
```

Components 9 - 10:
```{r}
as.data.frame(pca_output$var$coord[, 9:10]) %>% 
  mutate(stock = factor(row.names(pca_output$var$coord))) %>% 
  gather(key = "component", value = "coord", -stock) %>% 
  ggplot(aes(x = stock, y = coord, group = component, colour = component)) +
    geom_line(linewidth = 0.2) +  
    geom_point(size = 0.5) +
    theme(axis.text.x = element_blank())
```

Instead we could look at bi-plots...


### Bi-plot

Reference:  
https://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/112-pca-principal-component-analysis-essentials/#biplot 

Arrow colors represent cluster numbers from k-Sharp clustering.

Roughly speaking a bi-plot can be interpreted as follows:  

- An individual that is on the same side of a given variable has a high value for this variable;  
- An individual that is on the opposite side of a given variable has a low value for this variable.  



Dim 1+2
```{r}
fviz_pca_biplot(
    pca_output,
    axes = c(1,2),
    geom = "point",
    #col.ind = "contrib", ## color according to individual contribution to PC
    col.ind = "gray",
    col.var = factor(py$mcap_clusters), #"contrib", ## color according to contribution of variable to PC
    #gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    palette = "simpsons",
    label = "none"
  ) +
  labs(title = "Biplot of PCA Results")
```


```{r}
fviz_pca_var(
  pca_output, 
  axes = c(1,2),
  col.var = "contrib", #"black",
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  alpha.var = 0.25,
  label = "none"  
)
```




Dim 3+4
```{r}
fviz_pca_biplot(
    pca_output,
    axes = c(3,4),
    geom = "point",
    #col.ind = "contrib", ## color according to individual contribution to PC
    col.ind = "gray",
    col.var = factor(py$mcap_clusters), #"contrib", ## color according to contribution of variable to PC
    #gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    palette = "simpsons",
    label = "none"
  ) +
  labs(title = "Biplot of PCA Results")
```


```{r}
fviz_pca_var(
  pca_output, 
  axes = c(3,4),
  col.var = "contrib", #"black",
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  alpha.var = 0.25,
  label = "none"  
)
```


Dim 5+6
```{r}
fviz_pca_biplot(
    pca_output,
    axes = c(5,6),
    geom = "point",
    #col.ind = "contrib", ## color according to individual contribution to PC
    col.ind = "gray",
    col.var = factor(py$mcap_clusters), #"contrib", ## color according to contribution of variable to PC
    #gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    palette = "simpsons",
    label = "none"
  ) +
  labs(title = "Biplot of PCA Results")
```

```{r}
fviz_pca_var(
  pca_output, 
  axes = c(5,6),
  col.var = "contrib", #"black",
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  alpha.var = 0.25,
  label = "none"   
)
```


Dim 7+8
```{r}
fviz_pca_biplot(
    pca_output,
    axes = c(7,8),
    geom = "point",
    #col.ind = "contrib", ## color according to individual contribution to PC
    col.ind = "gray",
    col.var = factor(py$mcap_clusters), #"contrib", ## color according to contribution of variable to PC
    #gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    palette = "simpsons",
    label = "none"
  ) +
  labs(title = "Biplot of PCA Results")
```

```{r}
fviz_pca_var(
  pca_output, 
  axes = c(7,8),
  col.var = "contrib", #"black",
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  alpha.var = 0.25,
  label = "none"  
)
```


Dim 9+10
```{r}
fviz_pca_biplot(
    pca_output,
    axes = c(9,10),
    geom = "point",
    #col.ind = "contrib", ## color according to individual contribution to PC
    col.ind = "gray",
    col.var = factor(py$mcap_clusters), #"contrib", ## color according to contribution of variable to PC
    #gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    palette = "simpsons",
    label = "none"
  ) +
  labs(title = "Biplot of PCA Results")
```


```{r}
fviz_pca_var(
  pca_output, 
  axes = c(9,10),
  col.var = "contrib", #"black",
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  alpha.var = 0.25,
  label = "none" 
)
```

## Strategy ideas

### Idea 1

Grouping by clusters as an alternative to grouping by sector or industry.  

1) Group stocks into clusters using k-Sharp algorithm. Number of clusters is subject to budget constraints.  
2) From each cluster, select the stock closest to the centroid.  
3) Weight the stocks according to equal risk.  



### Idea 2

1) Each year, Apply PCA to the correlation matrix of a data set based on the previous two years of daily data.  
2) Exclude $PC_1$, as all stocks will have some degree of correlation with market risk ($PC_1$).  
3) Identify one stock for each used $PC_i$. Yang (see also Joliffe) offers this procedure:  
  1. Associate one variable with the highest coefficient in absolute value with each of the last $m_1$ principal components that have eigenvalue less than a certain level $l$ which we call the deletion criteria, then delete those $m_1$ variables. For example, one can use Kaiser’s rule. Recall that in the case of a correlation matrix, a principal component with eigenvalues smaller than 1 contains less information than one of the original variables.  
  2. A second PCA is performed on remaining variables. The same procedure was applied that associates one variable with each $m_2$ principal components that have an eigenvalue less than $l$, and delete those $m_2$ variables.  
  3. The procedure is repeated until no further deletions are considered necessary based on a stopping criteria. One can decide to stop the selection procedure based on the eigenvalue of the last principal component. For example, the stopping criteria can be delete variables until the retaining variables all have eigenvalue not less than 0.7.  
  - Alternative, simpler stock (variable) selection method (see Jolliffe, p 108):
    - Associate one variable with each of the first $m$ PCs, namely the variable not already chosen, with the highest coefficient, in absolute value, in each successive PC. These $m$ variables are retained, and the remaining $m^* = p - m$ are deleted. (This method is said to produce more "best" choices and more "bad" choices, while fewer "moderate" choices.)
4) Weight each stock $S_i$ by the ratio given by the stock's contribution to the corresponding $PC_i$ (given as contribution or coordinate) divided by the product of stock $S_i$'s contribution to $PC_1$ and the eigenvalue of $PC_i$, then normalized to sum to 1. (I.e. ignoring the stock's contribution to others PC's.)
  - The idea is balance the amount of variance contributed by each $PC_i|_{i>1}$ (each PC represented by one stock), and then adjust for each stock's contribution to the market risk represented by $PC_1$.  
5) Normalize the weighted portfolio to the desired volatility target.  


#### Implementation of idea 2

Assume that we are limited by budget constraints to 9 stocks. This implies including 10 PCs.

1) Each year, perform PCA based on the previous two years of daily data.  
```{r}
num_years_offset <- 0
pca_output <- PCA(top_mcap_returns_df_short[1:504 + (num_years_offset * 252), ], scale.unit = TRUE, ncp = 10, graph = FALSE)
```



3) Identify one stock for each used $PC_i$. Yang (see also Joliffe) offers this procedure:  
  1. For each $PC_j|_{1<j\leq m}$ select the variable (stock) with the highest absolute coordinate value:
```{r}
num_stocks <- ncol(pca_output$var$coord) - 1
coord_for_each_selected_stock <- numeric(num_stocks)
for(i in 1:num_stocks) {
  max_coord_i <- max(abs(pca_output$var$coord[ ,i+1]))
  coord_for_each_selected_stock[i] <- max_coord_i
  names(coord_for_each_selected_stock)[i] <- row.names(pca_output$var$coord)[which(abs(pca_output$var$coord[ ,i+1]) == max_coord_i)][1] ## Select the first if multiple stocks have exactly same coordinates
}
coord_for_each_selected_stock
```



4) Weight each stock $S_i$ by the ratio given by the stock's contribution to the corresponding $PC_i$ (given as absolute coordinate) divided by the product of stock $S_i$'s contribution to $PC_1$ and the eigenvalue of $PC_i$, then normalized to sum to 1. (I.e. ignoring the stock's contribution to others PC's.)
  - The idea is balance the amount of variance contributed by each $PC_i|_{i>1}$ (each PC represented by one stock), and then adjust for each stock's contribution to the market risk represented by $PC_1$.
  
```{r}
weights <- numeric(length(coord_for_each_selected_stock))
pc1_coords <- pca_output$var$coord[ ,1]
selected_symbols <- names(coord_for_each_selected_stock)
for(i in seq_along(coord_for_each_selected_stock)) {
  stock_coord <- coord_for_each_selected_stock[i]
  weight_in_pc1 <- pc1_coords[which(row.names(pca_output$var$coord) == selected_symbols[i])]
  eigenvalue <- pca_output$eig[i + 1, 1]
  weights[i] <- stock_coord / (weight_in_pc1 * eigenvalue)
}
weights <- weights / sum(weights)
names(weights) <- selected_symbols

weights
```

```{r}
selected_symbols_ids <- which(names(top_mcap_prices_df_short) %in% selected_symbols)
ggplot(
  aes(x = Index, y = Value, colour = Series), 
  ## OBS!
  ## fortify() converts time series to data frame.
  ## Use broom package instead!
  data = fortify(top_mcap_prices_df_short[, selected_symbols_ids], melt = TRUE)) + 
    geom_line(linewidth = 0.2) + 
    labs(title = "Prices of selected stocks, yr 1+2", x ="Time", y = "Price") +
    theme(legend.position="none")
```

```{r}
selected_symbols_ids <- which(names(top_mcap_prices_df_short) %in% selected_symbols)
ggplot(
  aes(x = Index, y = Value, colour = Series), 
  data = fortify(scale(top_mcap_prices_df_short[, selected_symbols_ids]), melt = TRUE)) + 
    geom_line(linewidth = 0.2) + 
    labs(title = "Scaled prices of selected stocks, yr 1+2", x ="Time", y = "Price") +
    theme(legend.position="none")
```

5) Normalize the weighted portfolio to the desired volatility target. 

TODO: Implement safe guards for risk weighting.
```{r}
sds <- lapply(top_mcap_prices_df_short[, selected_symbols_ids], function(x) {sd(x, na.rm = TRUE)})
#risk_normalized_weights <- lapply(sds, function(x) {x/sum(unlist(sds))})
risk_adjusted_weights <- unlist(lapply(sds, function(x) {1/x}))
risk_normalized_weights <- risk_adjusted_weights/sum(risk_adjusted_weights)

selected_pf_prices <- as.matrix(top_mcap_prices_df_short[, selected_symbols_ids]) %*% unname(risk_normalized_weights)

risk_normalized_weights
```

Plot weighted portfolio average of selected stocks.  
```{r}
selected_portfolio <- data.frame(
  date = index(top_mcap_prices_df_short),
  price = selected_pf_prices
)
  
  
ggplot(aes(x = date, y = price), data = selected_portfolio) + 
    geom_line(linewidth = 0.2) + 
    labs(title = "Avg. of selected portfolio", x ="Time", y = "Price") +
    theme(legend.position="none")
```


## Explained variation

We want to know the percentage of variance explained by the selected portfolio relative to the entire pool of candidate stocks.  
The variance of components is additive, as components are orthogonal by construction. If we assumed (unrealistically) that for each PC all stocks with a weight effectively larger than 0 were perfectly correlated, the selected stocks for each PC would then express same percentage of variation as it's associated PC. This scenario would be trivial, so let's ignore it.

Let's now consider the opposite extreme. Here we assume that all stocks affecting each PC were perfectly _un_correlated. This is of course again totally unrealistically. Recall, that the point of using PCA is exactly that the variables with dominating weights for each PC are more or less correlated, so that we can select one or a few variables for each PC to represent that dimension.  

TODO:  
How should we go about this? Adjust for covariances between stocks?  
Note here that for scaled data, correlations and covariances are the same.

For now let's do the calculations, given the assumption of uncorrelated stocks.

```{r}
pca_output <- PCA(top_mcap_returns_df_short[1:504, ], scale.unit = TRUE, ncp = 347)
```


Let  
$c_{i,j}$ be the contribution of stock $j$ to $PC_i$, and  
$v_i$ the percentage of variance expressed by $PC_i$, and  
$n$ the number of PC's.  
Then we calculate the percentage variance $U_j$ expressed by each stock as  
$$u_j = \sum_{i = 1}^n c_{i,j} v_i$$  

Percentages are decimalpercentages, i.e. $1.00$ rather than $100\%$.

```{r}
stock_contribution <- function(contribution_tbl, eigenvalue_tbl, stock_id, num_components) {
  c_j <- contribution_tbl[stock_id, ]/100
  v_ <- eigenvalue_tbl[, 2]/100 ## Column 2 in eigenvalue table is "percentage of variance" for each component
  u_j <- 0
  
  for(i in 1:num_components) {
    u_j <- u_j + c_j[i] * v_[i]
  }
  unname(u_j)
}
```

```{r}
stock_contribution(
  contribution_tbl = pca_output$var$contrib, 
  eigenvalue_tbl = pca_output$eig, 
  stock_id = 1, 
  num_components = 347
)
```



Then, for $m$ selected stocks, the percentage of variance, $u^*$, expressed by the selected portfolio of stocks is calculated as  
$$u^* =  \sum_{i = 1}^n \sum_{j = 1}^m c_{i,j} v_i$$

```{r}
pf_contribution <- function(contribution_tbl, eigenvalue_tbl, stock_id, num_components, pf_symbols, pf_weights) {
  u_ <- 0
  num_stocks <- length(pf_symbols)
  stock_pool <- row.names(contribution_tbl)
  stock_ids <- stock_pool[which(stock_pool %in% pf_symbols)]
  for(j in 1:num_stocks) {
    u_ <-  u_ + stock_contribution(
      contribution_tbl = pca_output$var$contrib, 
      eigenvalue_tbl = pca_output$eig, 
      stock_id = stock_ids[j], 
      num_components = num_components
    ) * pf_weights[j]
  }
  unname(u_)
}
```

```{r}
pf_contribution(
  contribution_tbl = pca_output$var$contrib, 
  eigenvalue_tbl = pca_output$eig, 
  stock_id = 1, 
  num_components = 347,
  pf_symbols = selected_symbols,
  pf_weights = risk_normalized_weights
)
```


## Systematic vs non-systematic risk

We want to know the ratio between systematic and non-systematic risk in the selected portfolio. Here we define systematic risk as that captured by $PC_1$. So for the selected portfolio we want to know the ratio between it's contributions to $PC_1$ and $\{PC_i\}_{i>1}$.

TODO:  
How do we go about this?

Could we use the following result from CAPM pf theory as an approximation?

$$\sigma_M = \sum_{j=1} w_{j,M} \sigma_{j, M}$$

where $\sigma_M$ is the market risk (volatility), and $\sigma_{j_M}$ is the covariance between the return of the $j$'th of $N$ risky assets in the market portfolio and the return of the entire market portfolio. The result implies that $w_{j,M} \sigma_{j_M}$ is the contribution of the $j$'th asset to the risk of the market portfolio.

See Rupert p. 505, section 17.5.1.

Alternatively, see section 17.6.1 *Estimation Using Regression* (p. 507).

Alternatively, see *Clifford S. Asness, Tobias J. Moskowitz, And Lasse Heje Pedersen: Value and Momentum Everywhere* on Eig\% (p. 46 in PDF).

Or simplest of all: Use the correlation between the 1st and the $j$'th PC.



## References

- Robert Carver: Smart Portfolios  
  - Complete framework  
  - Pragmatic  
  - Shows how things are done  
  - Many tables of worked out examples  
  - No code  
  - Minimal math  
  
- Rupert/Matteson: Statistics and Data Analysis for Financial Engineering 
  - Ch. 16, 17, 18, 20  
  - Good introduction to the mathematics of pf selection  
  - R code examples  
  
- Clifford S. Asness, Tobias J. Moskowitz, And Lasse Heje Pedersen: Value and Momentum Everywhere
  
  
  
