---
title: "Profiling 3"
author: "mhv"
date: "12/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Conclusion

+ Version 1:
  + Update the same element (here the last) in a large data frame many times.
+ Version 2:
  + Update a variable many times, then place it in an element in a large data 
  frame.
  
This tests the situation where a variable is assigned to different intermediate values before the final result is reached.
  
Version 2 is very much faster.

See also [/Users/mhvpbp13/Documents/Notes/Investering/rsystrade/Notes/rsystrade/Development/Coding paradigms - rsystrade.md](/Users/mhvpbp13/Documents/Notes/Investering/rsystrade/Notes/rsystrade/Development/Coding paradigms - rsystrade.md)

# Instructions:

Run everything at once.

# Performance test 3

```{r}
library("microbenchmark")
library("ggplot2")
library("magrittr")
```

## Setup

```{r}
nrows <- 252 * 25 ## 6300 (25 years of trade data for single table)
ncols <- 20
```


## Version 1
Update the same element (here the last) in a large data frame many times.

```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

version1 <- function(df) {
  for(i in 1:nrow(df)) {
    df[nrows, ncols] <- i
  }
  df
}

version1_out <- version1(df)
version1_out[1:2, ]

rm(df)
```

## Version 2
Update a variable many times, then place it in an element in a large data frame.

```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

version2 <- function(df) {
  var <- 0
  for(i in 1:nrow(df)) {
   var <- i
  }
  df[nrows, ncols] <- var
  df
}

version2_out <- version2(df)
version2_out[1:2, ]

rm(df)
```


```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

mb_out <- microbenchmark(
  version1(df),
  version2(df),
  times = 20L
)

mb_plot <- autoplot(mb_out) +
  geom_jitter(position = position_jitter(0.2, 0), 
              aes(color = expr), alpha = 0.4) + 
  aes(fill = I("gray")) + 
  theme(legend.position = "none") 
mb_plot
#labs(title = "Kernel smoother", subtitle = "")

rm(df)
```

```{r}
summary(mb_out)
```

Run benchmark test again
```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

mb_out <- microbenchmark(
  version1(df),
  version2(df),
  times = 20L
)
summary(mb_out)

rm(df)
```



```{r eval=FALSE, include=FALSE}
detach("package:microbenchmark", unload=TRUE)
detach("package:ggplot2", unload=TRUE)
detach("package:purrr", unload=TRUE)
detach("package:tibble", unload=TRUE)
detach("package:dplyr", unload=TRUE)
detach("package:magrittr", unload=TRUE)
```
