---
title: "Profiling 2"
author: "mhv"
date: "12/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Conclusion

+ Version 1:
  + Store all function outputs directly in a big data frame.
+ Version 2:
  + Store all function outputs in variables, then place them in the dataframe as
  a vector for each row.
+ Version 3:
  + Like version 1, but without the inner loop. Instead the assignments of the
  elements in each row are hard coded.
  + This checks that the difference between version 1 and version 2 is not just
  caused by the extra loop in version 1.

Version 2 is very much faster than both version 1 and version 3: It is better to
assign outputs to separate variables and then update each row in the data frame
in one step, rather than place the outs of functions directly into the df.

See also [/Users/mhvpbp13/Documents/Notes/Investering/rsystrade/Notes/rsystrade/Development/Coding paradigms - rsystrade.md](/Users/mhvpbp13/Documents/Notes/Investering/rsystrade/Notes/rsystrade/Development/Coding paradigms - rsystrade.md)

# Instructions:

Run everything at once.

# Performance test 2

```{r}
library("microbenchmark")
library("ggplot2")
library("magrittr")
```

## Setup

```{r}
#nrows <- 252 * 25 ## 6300 (25 years of trade data for single table)
nrows <- 252 ## One year of trade data
ncols <- 20
```


## Version 1
Store all function outputs directly in a big data frame.

```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

version1 <- function(df) {
  for(i in 1:nrow(df)) {
    for(j in 1:ncol(df)) {
      df[i, j] <- j
    }
  }
  df
}

version1_out <- version1(df)
version1_out[1:2, ]

rm(df)
```

## Version 2
Store all function outputs in variables, then place them in the dataframe as a
vector for each row.

```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

version2 <- function(df) {
  var1 <- 0
  var2 <- 0
  var3 <- 0
  var4 <- 0
  var5 <- 0
  var6 <- 0
  var7 <- 0
  var8 <- 0
  var9 <- 0
  var10 <- 0
  var11 <- 0
  var12 <- 0
  var13 <- 0
  var14 <- 0
  var15 <- 0
  var16 <- 0
  var17 <- 0
  var18 <- 0
  var19 <- 0
  var20 <- 0
  for(i in 1:nrow(df)) {
    var1 <- 1
    var2 <- 2
    var3 <- 3
    var4 <- 4
    var5 <- 5
    var6 <- 6
    var7 <- 7
    var8 <- 8
    var9 <- 9
    var10 <- 10
    var11 <- 11
    var12 <- 12
    var13 <- 13
    var14 <- 14
    var15 <- 15
    var16 <- 16
    var17 <- 17
    var18 <- 18
    var19 <- 19
    var20 <- 20
    df[i, ] <- c(
      var1,
      var2,
      var3,
      var4,
      var5,
      var6,
      var7,
      var8,
      var9,
      var10,
      var11,
      var12,
      var13,
      var14,
      var15,
      var16,
      var17,
      var18,
      var19,
      var20
    )
  }
  df
}

version2_out <- version2(df)
version2_out[1:2, ]

rm(df)
```

## Version 3
Store all function outputs directly in a big data frame.
Like version 1, but eliminate inner loop.

```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

version3 <- function(df) {
  for(i in 1:nrow(df)) {
      df[i, 1] <- 1
      df[i, 2] <- 2
      df[i, 3] <- 3 
      df[i, 4] <- 4 
      df[i, 5] <- 5
      df[i, 6] <- 6
      df[i, 7] <- 7
      df[i, 8] <- 8
      df[i, 9] <- 9
      df[i, 10] <- 10
      df[i, 11] <- 11
      df[i, 12] <- 12
      df[i, 13] <- 13
      df[i, 14] <- 14
      df[i, 15] <- 15
      df[i, 16] <- 16
      df[i, 17] <- 17
      df[i, 18] <- 18
      df[i, 19] <- 19
      df[i, 20] <- 20
  }
  df
}

version3_out <- version3(df)
version3_out[1:2, ]

rm(df)
```


```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

mb_out <- microbenchmark(
  version1(df),
  version2(df),
  version3(df),
  times = 20L
)

mb_plot <- autoplot(mb_out) +
  geom_jitter(position = position_jitter(0.2, 0), 
              aes(color = expr), alpha = 0.4) + 
  aes(fill = I("gray")) + 
  theme(legend.position = "none") 
mb_plot
#labs(title = "", subtitle = "")

rm(df)
```

```{r}
summary(mb_out)
```

Run benchmark test again
```{r}
df <- data.frame(matrix(rep(0, nrows * ncols ), ncol = ncols))

mb_out <- microbenchmark(
  version1(df),
  version2(df),
  version3(df),
  times = 20L
)
summary(mb_out)

rm(df)
```



```{r eval=FALSE, include=FALSE}
detach("package:microbenchmark", unload=TRUE)
detach("package:ggplot2", unload=TRUE)
detach("package:purrr", unload=TRUE)
detach("package:tibble", unload=TRUE)
detach("package:dplyr", unload=TRUE)
detach("package:magrittr", unload=TRUE)
```
